\chapter{Functional Analytic Foundations}\label{ch:foundations}

This chapter collects the functional analytic prerequisites for the Leray--Hopf theory
developed in \Cref{ch:leray-hopf}.
We build the framework in six steps: distributions and test functions
(\Cref{sec:distributions}), weak derivatives and Sobolev spaces (\Cref{sec:sobolev}),
Sobolev embedding theorems (\Cref{sec:embeddings}), the Rellich--Kondrachov compact
embedding theorem (\Cref{sec:rellich}), the Poincar\'{e} inequality (\Cref{sec:poincare}),
and the Helmholtz decomposition together with the Leray projector (\Cref{sec:helmholtz}).
The standard references are Evans~\cite{evans2010} and Adams--Fournier~\cite{adams2003};
for the Navier--Stokes setting see Temam~\cite{temam2001} and Robinson et al.\
\cite{robinson2016}.

Throughout, $\Omega \subset \R^{n}$ denotes an open set, and $n \ge 1$ unless otherwise
stated.
For the fluid applications in \Cref{ch:leray-hopf} we set $n = 3$.

% ---------------------------------------------------------------------------
\section{Distributions and test functions}\label{sec:distributions}
% ---------------------------------------------------------------------------

\subsection{Test functions and the LF topology}

\begin{definition}[Test functions]\label{def:test-functions}
  Let $K \subset \Omega$ be compact.  Define
  \[
    \mathcal{D}_{K}(\Omega)
    \;=\;
    \bigl\{\,\varphi \in C^{\infty}(\Omega) : \supp \varphi \subseteq K\,\bigr\}
  \]
  with the Fr\'{e}chet topology induced by the family of seminorms
  \[
    p_{m}(\varphi)
    \;=\;
    \max_{\abs{\alpha} \le m}
    \sup_{x \in \Omega} \abs{\partial^{\alpha}\varphi(x)},
    \qquad m \in \N.
  \]
  The \emph{space of test functions} is
  \[
    \mathcal{D}(\Omega)
    \;=\;
    \Ccinf(\Omega)
    \;=\;
    \bigcup_{K \Subset \Omega} \mathcal{D}_{K}(\Omega),
  \]
  equipped with the \emph{LF (strict inductive limit) topology}: a sequence
  $(\varphi_{j})$ converges in $\mathcal{D}(\Omega)$ if and only if there exists a
  compact set $K \Subset \Omega$ such that $\supp \varphi_{j} \subseteq K$ for all $j$
  and $p_{m}(\varphi_{j} - \varphi) \to 0$ for every $m \in \N$.
\end{definition}

\begin{remark}
  The LF topology makes $\mathcal{D}(\Omega)$ a complete locally convex space,
  though it is not metrisable.
  Sequential continuity nevertheless suffices for the theory of distributions.
\end{remark}

\subsection{Distributions}

\begin{definition}[Distributions]\label{def:distributions}
  A \emph{distribution} on $\Omega$ is a sequentially continuous linear functional
  $T : \mathcal{D}(\Omega) \to \R$.  The space of all distributions is the topological
  dual
  \[
    \mathcal{D}'(\Omega),
  \]
  equipped with the weak-$*$ topology.  We write $\langle T, \varphi \rangle$ for the
  duality pairing.
\end{definition}

\begin{example}[Locally integrable functions]\label{ex:L1loc}
  Every $f \in L^{1}_{\mathrm{loc}}(\Omega)$ defines a distribution $T_{f} \in
  \mathcal{D}'(\Omega)$ by
  \[
    \langle T_{f}, \varphi \rangle
    \;=\;
    \int_{\Omega} f(x)\,\varphi(x)\,\dd x,
    \qquad \varphi \in \mathcal{D}(\Omega).
  \]
  The assignment $f \mapsto T_{f}$ is injective, so $L^{1}_{\mathrm{loc}}(\Omega)$ embeds
  continuously into $\mathcal{D}'(\Omega)$.
\end{example}

\subsection{Distributional derivatives}

\begin{definition}[Distributional derivative]\label{def:distrib-deriv}
  For a multi-index $\alpha \in \N^{n}$ with $\abs{\alpha} = \alpha_{1} + \cdots +
  \alpha_{n}$, the \emph{distributional (partial) derivative} of $T \in
  \mathcal{D}'(\Omega)$ is the distribution $\partial^{\alpha} T$ defined by
  \[
    \langle \partial^{\alpha} T, \varphi \rangle
    \;=\;
    (-1)^{\abs{\alpha}}
    \langle T, \partial^{\alpha} \varphi \rangle,
    \qquad \varphi \in \mathcal{D}(\Omega).
  \]
  The sign is chosen so that, for smooth $f$, the distributional derivative $\partial^{\alpha}
  T_{f}$ coincides with $T_{\partial^{\alpha} f}$ (integration by parts).
\end{definition}

% ---------------------------------------------------------------------------
\section{Weak derivatives and Sobolev spaces}\label{sec:sobolev}
% ---------------------------------------------------------------------------

\subsection{Weak partial derivatives}

\begin{definition}[Weak partial derivative%
  \footnote{Lean 4: \texttt{NavierStokes.IsWeakPartialDeriv}}]%
  \label{def:weak-deriv}
  Let $u \in L^{1}_{\mathrm{loc}}(\Omega)$ and $\alpha \in \N^{n}$.  A function
  $v \in L^{1}_{\mathrm{loc}}(\Omega)$ is a \emph{weak $\alpha$-th partial derivative} of
  $u$, written $v = D^{\alpha} u$, if
  \begin{equation}\label{eq:weak-deriv}
    \int_{\Omega} v(x)\,\varphi(x)\,\dd x
    \;=\;
    (-1)^{\abs{\alpha}}
    \int_{\Omega} u(x)\,\partial^{\alpha}\varphi(x)\,\dd x
    \quad \text{for all } \varphi \in \Ccinf(\Omega).
  \end{equation}
\end{definition}

\begin{theorem}[Uniqueness of weak derivatives%
  \footnote{Lean 4: \texttt{NavierStokes.weakPartialDeriv\_unique}}]%
  \label{thm:weak-deriv-unique}
  If $v_{1}, v_{2} \in L^{1}_{\mathrm{loc}}(\Omega)$ are both weak $\alpha$-th partial
  derivatives of $u$, then $v_{1} = v_{2}$ almost everywhere in $\Omega$.
\end{theorem}

\begin{proof}
  Let $w = v_{1} - v_{2}$.  By definition,
  \[
    \int_{\Omega} w(x)\,\varphi(x)\,\dd x = 0
    \qquad \text{for all } \varphi \in \Ccinf(\Omega).
  \]
  This is precisely the content of the \emph{fundamental lemma of the calculus of
  variations}: if $w \in L^{1}_{\mathrm{loc}}(\Omega)$ satisfies $\int_{\Omega} w \varphi
  = 0$ for all $\varphi \in \Ccinf(\Omega)$, then $w = 0$ almost everywhere.  Indeed,
  for any Lebesgue point $x_{0}$ of $w$ and any sequence of mollifiers
  $\varphi_{\eps} = \rho_{\eps}(\cdot - x_{0})$ supported in $\Omega$, one obtains
  $w(x_{0}) = \lim_{\eps \to 0} \int w \varphi_{\eps} = 0$.
\end{proof}

\subsection{Sobolev spaces}

\begin{definition}[$W^{1,p}(\Omega)$%
  \footnote{Lean 4: \texttt{NavierStokes.SobolevW1p}}]%
  \label{def:W1p}
  For $1 \le p \le \infty$, the \emph{Sobolev space} $W^{1,p}(\Omega)$ consists of
  all $u \in \Lp{p}(\Omega)$ whose weak partial derivatives $D_{i} u$ exist in
  $\Lp{p}(\Omega)$ for each $i = 1, \ldots, n$.  It is a Banach space under
  \begin{equation}\label{eq:W1p-norm}
    \norm{u}_{W^{1,p}(\Omega)}
    \;=\;
    \begin{cases}
      \displaystyle
      \Bigl(\norm{u}_{\Lp{p}}^{p}
        + \sum_{i=1}^{n} \norm{D_{i} u}_{\Lp{p}}^{p}\Bigr)^{1/p},
      & 1 \le p < \infty, \\[8pt]
      \norm{u}_{L^{\infty}} + \displaystyle\sum_{i=1}^{n} \norm{D_{i} u}_{L^{\infty}},
      & p = \infty.
    \end{cases}
  \end{equation}
\end{definition}

\begin{definition}[$\Sob{1}(\Omega) = W^{1,2}(\Omega)$%
  \footnote{Lean 4: \texttt{NavierStokes.SobolevH1}}]%
  \label{def:H1}
  The space $\Sob{1}(\Omega) = W^{1,2}(\Omega)$ is a \emph{Hilbert space} with inner
  product
  \begin{equation}\label{eq:H1-inner-product}
    \ip{u}{v}_{\Sob{1}}
    \;=\;
    \int_{\Omega} u(x)\,v(x)\,\dd x
    \;+\;
    \sum_{i=1}^{n} \int_{\Omega} D_{i} u(x)\,D_{i} v(x)\,\dd x.
  \end{equation}
\end{definition}

\begin{definition}[$\Sobz{1}(\Omega)$%
  \footnote{Lean 4: \texttt{NavierStokes.SobolevH1Zero}}]%
  \label{def:H10}
  The space $\Sobz{1}(\Omega)$ is defined as the closure of $\Ccinf(\Omega)$ in
  $\Sob{1}(\Omega)$:
  \[
    \Sobz{1}(\Omega)
    \;=\;
    \overline{\Ccinf(\Omega)}^{\,\norm{\cdot}_{\Sob{1}}}.
  \]
  Functions in $\Sobz{1}(\Omega)$ satisfy a homogeneous Dirichlet boundary condition in
  the trace sense: $u\big|_{\partial\Omega} = 0$ whenever $\partial\Omega$ is
  sufficiently regular.
\end{definition}

% ---------------------------------------------------------------------------
\section{Sobolev embedding theorems}\label{sec:embeddings}
% ---------------------------------------------------------------------------

The embedding theorems quantify how membership in a Sobolev space controls pointwise
or $\Lp{q}$ regularity.  The critical exponent for $W^{1,p}(\R^{n})$ is
\[
  p^{*}
  \;=\;
  \frac{np}{n - p},
  \qquad 1 \le p < n.
\]

\subsection{Subcritical (Gagliardo--Nirenberg--Sobolev) embedding}

\begin{theorem}[Subcritical Sobolev embedding%
  \footnote{Lean 4: \texttt{NavierStokes.sobolev\_embedding\_subcritical}}]%
  \label{thm:sobolev-subcritical}
  Let $1 \le p < n$.  There exists a constant $C = C(n, p) > 0$ such that for all
  $u \in W^{1,p}(\R^{n})$,
  \begin{equation}\label{eq:sobolev-subcritical}
    \norm{u}_{\Lp{p^{*}}(\R^{n})}
    \;\le\;
    C\,\norm{\nabla u}_{\Lp{p}(\R^{n})}.
  \end{equation}
  Consequently, $W^{1,p}(\R^{n}) \hookrightarrow \Lp{p^{*}}(\R^{n})$ continuously.
\end{theorem}

\begin{proof}
  We first establish the case $p = 1$.  For $u \in \Ccinf(\R^{n})$ and each
  $i = 1, \ldots, n$, the fundamental theorem of calculus gives
  \[
    \abs{u(x)}
    \;\le\;
    \int_{-\infty}^{\infty}
    \abs{\partial_{i} u(x_{1}, \ldots, t, \ldots, x_{n})}\,\dd t,
  \]
  so
  \[
    \abs{u(x)}^{n/(n-1)}
    \;\le\;
    \prod_{i=1}^{n}
    \Bigl(\int_{-\infty}^{\infty} \abs{\partial_{i} u}\,\dd x_{i}\Bigr)^{1/(n-1)}.
  \]
  Integrating over $\R^{n}$ and applying the generalised H\"{o}lder inequality
  (with $n - 1$ factors, each in $\Lp{1}$ of the remaining variable) yields
  $\norm{u}_{\Lp{n/(n-1)}} \le \prod_{i} \norm{\partial_{i} u}_{\Lp{1}}^{1/n}
  \le \norm{\nabla u}_{\Lp{1}}$, which is the Gagliardo--Nirenberg inequality
  (the case $p = 1$, $p^{*} = n/(n-1)$).

  For $1 < p < n$, set $\gamma = p(n-1)/(n-p) > 1$ and apply the $p = 1$ result to
  $v = \abs{u}^{\gamma}$.  Since $\abs{\nabla v} = \gamma \abs{u}^{\gamma - 1}
  \abs{\nabla u}$, H\"{o}lder's inequality with exponents $p$ and $p/(p-1)$ gives
  $\norm{\nabla v}_{\Lp{1}} \le \gamma \norm{u}_{\Lp{\gamma p/(p-1)}}^{\gamma - 1}
  \norm{\nabla u}_{\Lp{p}}$.  The exponents are arranged so that $\gamma n/(n-1)
  = \gamma p/(p-1) = p^{*}$, and rearranging produces the desired inequality
  $\norm{u}_{\Lp{p^{*}}} \le C(n,p)\,\norm{\nabla u}_{\Lp{p}}$.

  The extension from $\Ccinf(\R^{n})$ to $W^{1,p}(\R^{n})$ follows by density:
  approximate $u \in W^{1,p}$ by a sequence $(u_{k}) \subset \Ccinf$ in the $W^{1,p}$
  norm, apply the inequality to each $u_{k}$, and pass to the limit using completeness
  of $\Lp{p^{*}}$.  For bounded domains with Lipschitz boundary, the Stein extension
  operator provides a bounded map $W^{1,p}(\Omega) \to W^{1,p}(\R^{n})$, reducing to
  the whole-space case.  See \cite{evans2010}, Section~5.6.
\end{proof}

\begin{remark}[The case $n = 3$, $p = 2$]
  For the Navier--Stokes problem with $n = 3$ and $p = 2$, one has
  \[
    p^{*} = \frac{3 \cdot 2}{3 - 2} = 6,
  \]
  so $\Sob{1}(\R^{3}) \hookrightarrow \Lp{6}(\R^{3})$ with the explicit bound
  \begin{equation}\label{eq:H1-L6-embedding}
    \norm{u}_{\Lp{6}(\R^{3})}
    \;\le\;
    C\,\norm{\nabla u}_{\Lp{2}(\R^{3})}.
  \end{equation}
  This embedding is used repeatedly in the energy estimates of \Cref{ch:leray-hopf}.
\end{remark}

\subsection{Supercritical (Morrey) embedding}

\begin{theorem}[Morrey embedding%
  \footnote{Lean 4: \texttt{NavierStokes.sobolev\_embedding\_supercritical}}]%
  \label{thm:sobolev-morrey}
  Let $p > n$ and set $\alpha = 1 - n/p \in (0, 1)$.  Then there exists a constant
  $C = C(n, p) > 0$ such that for all $u \in W^{1,p}(\R^{n})$,
  \begin{equation}\label{eq:morrey}
    \abs{u(x) - u(y)}
    \;\le\;
    C\,\abs{x - y}^{\alpha}\,\norm{u}_{W^{1,p}(\R^{n})},
    \qquad \text{a.e.\ } x, y \in \R^{n}.
  \end{equation}
  In particular, $W^{1,p}(\R^{n}) \hookrightarrow C^{0,\alpha}(\R^{n})$ continuously.
\end{theorem}

\begin{proof}
  By density it suffices to prove the estimate for $u \in \Ccinf(\R^{n})$.  Fix
  $x, y \in \R^{n}$ and set $r = \abs{x - y}$.  For any $z \in B_{r}(x)$, the
  fundamental theorem of calculus along the segment from $z$ to $x$ gives
  \[
    u(x) - u(z)
    \;=\;
    \int_{0}^{1} \nabla u\bigl(z + t(x - z)\bigr) \cdot (x - z)\,\dd t.
  \]
  Averaging over $z \in B_{r}(x)$ and denoting $u_{B_{r}} = \fint_{B_{r}(x)} u\,\dd z$,
  we obtain
  \begin{align*}
    \abs{u(x) - u_{B_{r}}}
    &\;\le\;
    \fint_{B_{r}(x)} \int_{0}^{1}
      \abs{\nabla u\bigl(z + t(x-z)\bigr)}\,\abs{x - z}\,\dd t\,\dd z \\
    &\;\le\;
    \frac{1}{\abs{B_{r}}}
    \int_{B_{r}(x)} \frac{\abs{\nabla u(w)}}{\abs{x - w}^{n-1}}\,\dd w,
  \end{align*}
  where the last inequality follows from the change of variables $w = z + t(x - z)$ and
  the bound $\abs{x - z} \le r$.  Applying H\"{o}lder's inequality with exponents $p$
  and $p' = p/(p-1)$ to the integral on the right yields
  \[
    \abs{u(x) - u_{B_{r}}}
    \;\le\;
    \frac{1}{\abs{B_{r}}}
    \norm{\nabla u}_{\Lp{p}(B_{r})}
    \Bigl(\int_{B_{r}(x)} \abs{x - w}^{-(n-1)p'}\,\dd w\Bigr)^{1/p'}.
  \]
  Since $p > n$, one has $(n-1)p' < n$, so the integral converges and equals
  $C(n,p)\,r^{n - (n-1)p'}$.  Combining this with $\abs{B_{r}} = \omega_{n}\,r^{n}$
  gives $\abs{u(x) - u_{B_{r}}} \le C\,r^{1 - n/p}\,\norm{\nabla u}_{\Lp{p}(B_{r})}$.
  The same bound holds with $x$ replaced by $y$ (since $B_{r}(y) \subset B_{2r}(x)$),
  so the triangle inequality yields
  \[
    \abs{u(x) - u(y)}
    \;\le\;
    \abs{u(x) - u_{B_{r}}} + \abs{u(y) - u_{B_{r}}}
    \;\le\;
    C\,\abs{x - y}^{1 - n/p}\,\norm{\nabla u}_{\Lp{p}(\R^{n})}.
  \]
  Setting $\alpha = 1 - n/p$ and absorbing the full $W^{1,p}$ norm via the Sobolev
  inequality completes the proof.  The extension from $\Ccinf$ to $W^{1,p}$ follows by
  approximation: the H\"{o}lder estimate passes to the limit, and the representative
  obtained is continuous after modification on a set of measure zero.
\end{proof}

% ---------------------------------------------------------------------------
\section{Compact embeddings: Rellich--Kondrachov}\label{sec:rellich}
% ---------------------------------------------------------------------------

The Rellich--Kondrachov theorem upgrades the continuous Sobolev embedding to a
\emph{compact} one, at the cost of passing to a lower-order space.
Compactness is essential for the convergence of Galerkin approximations in
\Cref{ch:leray-hopf}.

\begin{theorem}[Rellich--Kondrachov%
  \footnote{Lean 4: \texttt{NavierStokes.rellich\_kondrachov}}]%
  \label{thm:rellich-kondrachov}
  Let $\Omega \subset \R^{n}$ be a bounded open set with Lipschitz boundary.  Then the
  inclusion
  \[
    \Sob{1}(\Omega) \hookrightarrow \Lp{2}(\Omega)
  \]
  is \emph{compact}: every bounded sequence in $\Sob{1}(\Omega)$ has a subsequence that
  converges strongly in $\Lp{2}(\Omega)$.  More generally, for $1 \le p < n$ and
  $1 \le q < p^{*}$, the inclusion $W^{1,p}(\Omega) \hookrightarrow \Lp{q}(\Omega)$ is
  compact.
\end{theorem}

\begin{proof}
  Let $(u_{j})$ be bounded in $\Sob{1}(\Omega)$ with
  $\norm{u_{j}}_{\Sob{1}} \le M$ for all $j$.  We establish precompactness in
  $\Lp{2}(\Omega)$ by a mollification and diagonal argument.

  Let $\tilde{u}_{j}$ denote the zero extension of $u_{j}$ to $\R^{n}$, and let
  $\rho_{\eps}$ be a standard mollifier with $\supp \rho_{\eps} \subset B_{\eps}(0)$.
  Define $u_{j}^{\eps} = \rho_{\eps} * \tilde{u}_{j}$.  By the standard mollification
  estimate (a consequence of Jensen's inequality and the characterisation of $\Sob{1}$
  via difference quotients),
  \[
    \norm{u_{j}^{\eps} - u_{j}}_{\Lp{2}(\Omega)}
    \;\le\;
    \eps\,\norm{\nabla u_{j}}_{\Lp{2}(\Omega)}
    \;\le\;
    M\eps
  \]
  for every $j$ and every $\eps > 0$.

  For each fixed $\eps$, the mollified family $(u_{j}^{\eps})_{j}$ is uniformly bounded
  and equicontinuous on $\overline{\Omega}$.  Indeed, Young's convolution inequality gives
  \[
    \norm{u_{j}^{\eps}}_{\Lp{\infty}}
    \;\le\;
    \norm{\rho_{\eps}}_{\Lp{2}}\,\norm{u_{j}}_{\Lp{2}}
    \;\le\;
    C(\eps)\,M,
  \]
  and similarly
  $\norm{\nabla u_{j}^{\eps}}_{\Lp{\infty}} \le
  \norm{\nabla \rho_{\eps}}_{\Lp{2}}\,\norm{u_{j}}_{\Lp{2}} \le C'(\eps)\,M$, so
  \[
    \abs{u_{j}^{\eps}(x) - u_{j}^{\eps}(y)}
    \;\le\;
    C'(\eps)\,M\,\abs{x - y}
  \]
  for all $x, y \in \overline{\Omega}$.  By the Arzel\`{a}--Ascoli theorem, $(u_{j}^{\eps})_{j}$
  is precompact in $C(\overline{\Omega})$, hence also in $\Lp{2}(\Omega)$ (since $\Omega$
  is bounded).

  Now apply a diagonal argument.  Choose $\eps_{k} = 1/k$ and extract successive
  subsequences: from $(u_{j})$ extract $(u_{j}^{(1)})$ such that
  $(u_{j}^{(1),\,1/1})_{j}$ converges in $\Lp{2}$; from $(u_{j}^{(1)})$ extract
  $(u_{j}^{(2)})$ such that $(u_{j}^{(2),\,1/2})_{j}$ converges; and so on.  The
  diagonal sequence $v_{k} = u_{k}^{(k)}$ then satisfies the following: for any
  $\delta > 0$, choose $\eps$ so that $M\eps < \delta/3$, then choose $N$ so that
  $\norm{v_{j}^{\eps} - v_{k}^{\eps}}_{\Lp{2}} < \delta/3$ for all $j, k \ge N$.  The
  triangle inequality
  \[
    \norm{v_{j} - v_{k}}_{\Lp{2}}
    \;\le\;
    \norm{v_{j} - v_{j}^{\eps}}_{\Lp{2}}
    + \norm{v_{j}^{\eps} - v_{k}^{\eps}}_{\Lp{2}}
    + \norm{v_{k}^{\eps} - v_{k}}_{\Lp{2}}
    \;<\;
    \delta
  \]
  shows that $(v_{k})$ is Cauchy in $\Lp{2}(\Omega)$, and completeness gives strong
  convergence.  The general case $W^{1,p} \hookrightarrow \Lp{q}$ for $q < p^{*}$ follows
  by interpolation between the $\Lp{p^{*}}$ bound (from \Cref{thm:sobolev-subcritical})
  and the $\Lp{2}$ compactness just established.
\end{proof}

\begin{remark}[Role in the Galerkin method]
  In \Cref{ch:leray-hopf}, the Galerkin approximations $(u_{m})$ are constructed
  as bounded sequences in the energy class $L^{\infty}(0,T;\Lp{2}) \cap
  L^{2}(0,T;\Sob{1})$.  The Rellich--Kondrachov theorem (together with the Aubin--Lions
  lemma) provides the strong $\Lp{2}$ convergence needed to pass nonlinear terms to the
  limit.
\end{remark}

% ---------------------------------------------------------------------------
\section{The Poincar\'{e} inequality}\label{sec:poincare}
% ---------------------------------------------------------------------------

\begin{theorem}[Poincar\'{e} inequality%
  \footnote{Lean 4: \texttt{NavierStokes.poincare\_inequality}}]%
  \label{thm:poincare}
  Let $\Omega \subset \R^{n}$ be a bounded open set.  There exists a constant
  $C_{\mathrm{P}} = C_{\mathrm{P}}(\Omega) > 0$ such that
  \begin{equation}\label{eq:poincare}
    \norm{u}_{\Lp{2}(\Omega)}
    \;\le\;
    C_{\mathrm{P}}\,\norm{\nabla u}_{\Lp{2}(\Omega)}
    \qquad \text{for all } u \in \Sobz{1}(\Omega).
  \end{equation}
\end{theorem}

\begin{proof}
  Suppose for contradiction that no such constant exists.  Then there is a sequence
  $(u_{j}) \subset \Sobz{1}(\Omega)$ with
  \[
    \norm{u_{j}}_{\Lp{2}} = 1
    \qquad \text{and} \qquad
    \norm{\nabla u_{j}}_{\Lp{2}} \to 0.
  \]
  The sequence is bounded in $\Sob{1}(\Omega)$ (since $\norm{u_{j}}_{\Sob{1}}^{2} =
  \norm{u_{j}}_{\Lp{2}}^{2} + \norm{\nabla u_{j}}_{\Lp{2}}^{2} \le 1 + 1 = 2$), so by
  Rellich--Kondrachov (\Cref{thm:rellich-kondrachov}), there exists a subsequence
  (still denoted $(u_{j})$) with $u_{j} \to u$ strongly in $\Lp{2}(\Omega)$.
  In particular $\norm{u}_{\Lp{2}} = 1$.

  On the other hand, $\nabla u_{j} \to 0$ in $\Lp{2}(\Omega)$, and weak lower
  semicontinuity of the norm gives $\nabla u = 0$ almost everywhere.
  Since $\Omega$ is connected, $u$ is constant almost everywhere; since $u \in
  \Sobz{1}(\Omega)$, the trace of $u$ on $\partial\Omega$ is zero, so $u \equiv 0$.
  This contradicts $\norm{u}_{\Lp{2}} = 1$.
\end{proof}

\begin{corollary}[Equivalent norm on $\Sobz{1}(\Omega)$]\label{cor:poincare-norm}
  On $\Sobz{1}(\Omega)$, the seminorm $u \mapsto \norm{\nabla u}_{\Lp{2}(\Omega)}$ is a
  norm equivalent to $\norm{\cdot}_{\Sob{1}(\Omega)}$.  Concretely,
  \begin{equation}\label{eq:poincare-equivalence}
    \norm{\nabla u}_{\Lp{2}}
    \;\le\;
    \norm{u}_{\Sob{1}}
    \;\le\;
    \sqrt{1 + C_{\mathrm{P}}^{2}}\,\norm{\nabla u}_{\Lp{2}}.
  \end{equation}
  In particular, $(\Sobz{1}(\Omega),\, \ip{\nabla \cdot}{\nabla \cdot}_{\Lp{2}})$ is a
  Hilbert space.
\end{corollary}

% ---------------------------------------------------------------------------
\section[Divergence-free fields and Helmholtz]{Divergence-free fields and the Helmholtz decomposition}%
\label{sec:helmholtz}
% ---------------------------------------------------------------------------

The incompressibility constraint $\divergence u = 0$ is encoded variationally via the
distributional divergence-free condition.

\subsection{Distributional divergence-free fields}

\begin{definition}[Divergence-free fields]%
  \label{def:distrib-divfree}%
  \footnote{Lean 4: \texttt{NavierStokes.IsDistribDivFree}.}
  A vector field $u \in \Lp{2}(\Omega;\R^{n})$ is \emph{distributionally divergence-free}
  if
  \begin{equation}\label{eq:distrib-divfree}
    \int_{\Omega} u \cdot \nabla \varphi \,\dd x = 0
    \qquad \text{for all } \varphi \in \Ccinf(\Omega).
  \end{equation}
  Equivalently, $\langle \divergence T_{u},\, \varphi \rangle = 0$ for all $\varphi \in
  \Ccinf(\Omega)$, i.e., $\divergence u = 0$ in $\mathcal{D}'(\Omega)$.
\end{definition}

\subsection{The space $\Lsig(\Omega)$}

\begin{definition}[$\Lsig(\Omega)$%
  \footnote{Lean 4: \texttt{NavierStokes.L2sigma}}]%
  \label{def:L2sigma}
  Define
  \[
    \Lsig(\Omega)
    \;=\;
    \overline{
      \bigl\{\,
        u \in \Ccinf(\Omega;\R^{n}) : \divergence u = 0
      \,\bigr\}
    }^{\,\norm{\cdot}_{\Lp{2}}}.
  \]
  This is the $\Lp{2}$-closure of smooth, compactly supported, divergence-free vector
  fields.
\end{definition}

\begin{proposition}[Closedness of $\Lsig$]%
  \label{prop:L2sigma-closed}%
  \footnote{Lean 4: \texttt{NavierStokes.l2sigma\_isClosed}.}
  The space $\Lsig(\Omega)$ is a closed subspace of $\Lp{2}(\Omega;\R^{n})$, hence
  itself a Hilbert space.
\end{proposition}

\begin{proof}
  $\Lsig(\Omega)$ is defined as a closure, so it is closed by construction.
  Alternatively: the distributional divergence operator
  $\divergence : \Lp{2}(\Omega;\R^{n}) \to H^{-1}(\Omega)$ is continuous, so its kernel
  $\{u : \divergence u = 0\}$ is closed.  One verifies that this kernel equals the
  closure of the smooth divergence-free fields when $\Omega$ has the appropriate
  topology (e.g., $\Omega = \R^{n}$ or $\Omega$ simply connected with Lipschitz
  boundary).
\end{proof}

\subsection{Helmholtz decomposition}

\begin{theorem}[Helmholtz decomposition%
  \footnote{Lean 4: \texttt{NavierStokes.helmholtz\_decomposition}}]%
  \label{thm:helmholtz}
  Let $\Omega \subset \R^{n}$ be a simply connected bounded open set with $C^{1}$
  boundary, or $\Omega = \R^{n}$.  Then $\Lp{2}(\Omega;\R^{n})$ admits the orthogonal
  decomposition
  \begin{equation}\label{eq:helmholtz}
    \Lp{2}(\Omega;\R^{n})
    \;=\;
    \Lsig(\Omega)
    \;\oplus_{\perp}\;
    G(\Omega),
  \end{equation}
  where
  \[
    G(\Omega)
    \;=\;
    \bigl\{\,
      \nabla p : p \in \Sob{1}(\Omega)
    \,\bigr\}.
  \]
  Every $u \in \Lp{2}(\Omega;\R^{n})$ decomposes uniquely as
  $u = v + \nabla p$ with $v \in \Lsig(\Omega)$ and $\nabla p \in G(\Omega)$.
\end{theorem}

\begin{proof}
  We first establish orthogonality of $\Lsig(\Omega)$ and $G(\Omega)$.  Let $v \in
  \Lsig(\Omega)$ and $\nabla p \in G(\Omega)$.  Since $v$ is the $\Lp{2}$-limit of
  smooth divergence-free fields $v_{k} \in \Ccinf(\Omega;\R^{n})$, integration by parts
  gives $\ip{v_{k}}{\nabla p}_{\Lp{2}} = -\int_{\Omega} (\divergence v_{k})\,p\,\dd x
  = 0$ for each $k$, and passing to the limit yields $\ip{v}{\nabla p}_{\Lp{2}} = 0$.
  Hence $\Lsig(\Omega) \perp G(\Omega)$.

  To construct the decomposition, let $u \in \Lp{2}(\Omega;\R^{n})$ be given.  Consider
  the bilinear form $a : \Sob{1}(\Omega)/\R \times \Sob{1}(\Omega)/\R \to \R$ defined by
  $a(p, q) = \int_{\Omega} \nabla p \cdot \nabla q\,\dd x$ and the linear functional
  $\ell(q) = \int_{\Omega} u \cdot \nabla q\,\dd x$.  The form $a$ is continuous and
  coercive on $\Sob{1}(\Omega)/\R$ (coercivity follows from the Poincar\'{e}--Wirtinger
  inequality for functions with zero mean), and $\ell$ is continuous since
  $\abs{\ell(q)} \le \norm{u}_{\Lp{2}}\norm{\nabla q}_{\Lp{2}}$.  By the Lax--Milgram
  theorem, there exists a unique $p \in \Sob{1}(\Omega)/\R$ such that
  $a(p, q) = \ell(q)$ for all $q$, which is the weak formulation of the Neumann problem
  $-\Delta p = -\divergence u$ in $\Omega$ with $\partial p / \partial \nu = u \cdot \nu$
  on $\partial\Omega$.

  Set $v = u - \nabla p$.  For any $\varphi \in \Ccinf(\Omega)$, the weak formulation
  gives $\int_{\Omega} v \cdot \nabla \varphi\,\dd x = \int_{\Omega} u \cdot \nabla
  \varphi\,\dd x - \int_{\Omega} \nabla p \cdot \nabla \varphi\,\dd x = \ell(\varphi)
  - a(p, \varphi) = 0$, so $\divergence v = 0$ in $\mathcal{D}'(\Omega)$.  To verify
  that $v \in \Lsig(\Omega)$ (the $\Lp{2}$-closure of smooth divergence-free fields,
  not merely the distributional kernel), one approximates $v$ by mollification:
  $v_{\eps} = \rho_{\eps} * \tilde{v}$ satisfies $\divergence v_{\eps} = \rho_{\eps} *
  \divergence \tilde{v} = 0$ on any compact subset of $\Omega$ for $\eps$ sufficiently
  small, and $v_{\eps} \to v$ in $\Lp{2}$.  Multiplying by a smooth cutoff confirms
  $v \in \Lsig(\Omega)$.

  Uniqueness of the decomposition follows from the orthogonality already established: if
  $u = v_{1} + \nabla p_{1} = v_{2} + \nabla p_{2}$ with $v_{i} \in \Lsig(\Omega)$ and
  $\nabla p_{i} \in G(\Omega)$, then $v_{1} - v_{2} = \nabla(p_{2} - p_{1}) \in
  \Lsig(\Omega) \cap G(\Omega)$, and orthogonality forces $v_{1} - v_{2} = 0$.  On
  simply connected domains (or $\R^{n}$), the de Rham theorem ensures that every
  curl-free $\Lp{2}$ vector field is a gradient, so $G(\Omega)$ coincides with the
  orthogonal complement $\Lsig(\Omega)^{\perp}$, completing the proof.
\end{proof}

\subsection{The Leray projector}

\begin{definition}[Leray projector%
  \footnote{Lean 4: \texttt{NavierStokes.lerayProjector}}]%
  \label{def:leray-projector}
  The \emph{Leray (or Helmholtz--Leray) projector} is the bounded linear operator
  \[
    \Leray : \Lp{2}(\Omega;\R^{n}) \longrightarrow \Lsig(\Omega)
  \]
  defined by $\Leray u = v$, where $u = v + \nabla p$ is the Helmholtz decomposition of
  $u$.  Equivalently, $\Leray$ is the orthogonal projection onto $\Lsig(\Omega)$.
\end{definition}

\begin{proposition}[Properties of the Leray projector]\label{prop:leray-props}
  The operator $\Leray$ satisfies:
  \begin{enumerate}[label=(\roman*)]
    \item \emph{Boundedness.} $\norm{\Leray u}_{\Lp{2}} \le \norm{u}_{\Lp{2}}$ for all
      $u \in \Lp{2}(\Omega;\R^{n})$.
    \item \emph{Idempotency.} $\Leray^{2} = \Leray$.
    \item \emph{Self-adjointness.} $\ip{\Leray u}{v}_{\Lp{2}} = \ip{u}{\Leray
      v}_{\Lp{2}}$ for all $u, v \in \Lp{2}(\Omega;\R^{n})$.
    \item \emph{Projection onto divergence-free fields.} $\divergence(\Leray u) = 0$ for
      all $u$.
  \end{enumerate}
\end{proposition}

\begin{proof}
  Properties (i)--(iii) are standard consequences of orthogonal projection in Hilbert
  space.  Property (iv) holds by definition of $\Lsig(\Omega)$.
\end{proof}

\begin{remark}[Role in the Navier--Stokes equations]
  Applying $\Leray$ to the momentum equation
  $\partial_{t} u + (u \cdot \nabla) u - \nu \Delta u + \nabla p = f$ eliminates the
  pressure gradient $\nabla p$ (which lies in $G(\Omega)$) and yields the projected
  equation
  \begin{equation}\label{eq:projected-ns}
    \partial_{t} u + \Leray\bigl[(u \cdot \nabla) u\bigr] - \nu \Delta u = \Leray f.
  \end{equation}
  This is the form amenable to semigroup theory and the Galerkin method of
  \Cref{ch:leray-hopf}.
\end{remark}

\bigskip

\noindent\textbf{Summary.}
The tools assembled in this chapter form the complete functional analytic toolkit for
the weak theory of the Navier--Stokes equations.
The Sobolev space $\Sobz{1}(\Omega)$ provides the energy space for the velocity field;
the subcritical embedding $\Sob{1} \hookrightarrow \Lp{6}$ (for $n = 3$) controls the
nonlinear term; Rellich--Kondrachov furnishes the compactness needed in Galerkin limits;
the Poincar\'{e} inequality ensures coercivity; and the Helmholtz decomposition decouples
velocity from pressure via the Leray projector $\Leray$.
\Cref{ch:leray-hopf} combines these ingredients to construct Leray--Hopf weak
solutions.
