\chapter{Leray--Hopf Weak Solutions}\label{ch:leray-hopf}

In 1934, Jean Leray~\cite{leray1934} proved that the three-dimensional incompressible
Navier--Stokes equations admit global-in-time weak solutions for arbitrary $\Lp{2}$
initial data, and that these solutions satisfy a fundamental energy inequality.
Two decades later, Eberhard Hopf~\cite{hopf1951} extended the construction to bounded
domains.  The solutions produced by their method, now called \emph{Leray--Hopf weak
solutions}, remain the only class of weak solutions known to exist globally in three
dimensions.  Whether every Leray--Hopf weak solution is in fact smooth is the central
open question of the Clay Millennium Problem~\cite{fefferman2006}.

This chapter constructs Leray--Hopf weak solutions on $\R^{3}$ in full
detail.  We proceed from the strong formulation (\Cref{sec:ns-strong})
and its weak counterpart (\Cref{sec:ns-weak}) to the existence proof
via the Galerkin method.  \Cref{sec:clay} states the Clay Millennium
Problem.

Throughout this chapter, the spatial dimension is $n = 3$.  We use freely
the functional analytic tools from \Cref{ch:foundations}: Sobolev spaces,
embedding theorems, the Rellich--Kondrachov theorem, the Poincar\'{e}
inequality, and the Helmholtz decomposition with the Leray projector.

% ---------------------------------------------------------------------------
\section{The incompressible Navier--Stokes equations}\label{sec:ns-strong}
% ---------------------------------------------------------------------------

\subsection{Strong formulation}

The motion of an incompressible viscous fluid with unit density occupying $\R^{3}$ is
governed by the Navier--Stokes system: find a velocity field
$u : \R^{3} \times [0, \infty) \to \R^{3}$ and a pressure field
$p : \R^{3} \times [0, \infty) \to \R$ satisfying

\begin{equation}\label{eq:ns-momentum}
  \partial_{t} u + (u \cdot \nabla) u
  \;=\;
  \nu \Delta u - \nabla p + f,
  \qquad x \in \R^{3},\; t > 0,
\end{equation}
\begin{equation}\label{eq:ns-incompressibility}
  \divergence u
  \;=\;
  0,
  \qquad x \in \R^{3},\; t > 0,
\end{equation}
\begin{equation}\label{eq:ns-initial}
  u(x, 0)
  \;=\;
  u_{0}(x),
  \qquad x \in \R^{3},
\end{equation}
together with the decay condition $u(x, t) \to 0$ as $\abs{x} \to \infty$.

\begin{remark}[Physical meaning of each term]
  In the momentum equation~\eqref{eq:ns-momentum}:
  \begin{enumerate}[label=(\roman*)]
    \item $\partial_{t} u$ is the local (Eulerian) acceleration of the fluid;
    \item $(u \cdot \nabla) u = \sum_{j=1}^{3} u_{j}\,\partial_{j} u$ is the
      \emph{convective} (or \emph{advective}) nonlinearity, encoding the transport of
      momentum by the flow itself;
    \item $\nu \Delta u$ is the viscous diffusion term, with $\nu > 0$ the kinematic
      viscosity;
    \item $\nabla p$ is the pressure gradient, which enforces the incompressibility
      constraint~\eqref{eq:ns-incompressibility} (the pressure acts as a Lagrange
      multiplier);
    \item $f = f(x, t)$ is an external body force.
  \end{enumerate}
  The constraint $\divergence u = 0$ expresses conservation of mass for a fluid of
  constant density: the velocity field is volume-preserving.
\end{remark}

\subsection{The convective term and its symmetry}

The convective term $(u \cdot \nabla)u$ is the source of the mathematical difficulty.
Written in component form, it reads
\[
  \bigl[(u \cdot \nabla) u\bigr]_{i}
  \;=\;
  \sum_{j=1}^{3} u_{j}\,\partial_{j} u_{i},
  \qquad i = 1, 2, 3.
\]
When $u$ is divergence-free, this can be rewritten in conservative form:
\begin{equation}\label{eq:convective-conservative}
  (u \cdot \nabla) u
  \;=\;
  \divergence(u \otimes u),
\end{equation}
since $\divergence(u \otimes u)_{i} = \sum_{j} \partial_{j}(u_{j} u_{i})
= \sum_{j} u_{j}\,\partial_{j} u_{i} + u_{i}\,\underbrace{\divergence u}_{=\,0}$.
The conservative form~\eqref{eq:convective-conservative} is better suited to the weak
formulation because it requires only one derivative, distributed via integration by
parts.

\subsection{Dimensionless form and the Reynolds number}

Introduce characteristic scales: a length $L$, a velocity $U$, and a time $T = L/U$.
The dimensionless variables $x' = x/L$, $t' = t/T$, $u' = u/U$, $p' = p/(U^{2})$,
$f' = fL/U^{2}$ transform the Navier--Stokes equations into
\begin{equation}\label{eq:ns-dimensionless}
  \partial_{t'} u' + (u' \cdot \nabla') u'
  \;=\;
  \frac{1}{\mathrm{Re}}\,\Delta' u' - \nabla' p' + f',
\end{equation}
where
\begin{equation}\label{eq:reynolds}
  \mathrm{Re}
  \;=\;
  \frac{UL}{\nu}
\end{equation}
is the \emph{Reynolds number}.  The Reynolds number measures the ratio of inertial forces
(the convective term) to viscous forces (the diffusion term).  At low Reynolds number
($\mathrm{Re} \ll 1$), viscosity dominates and the flow is laminar; at high Reynolds
number ($\mathrm{Re} \gg 1$), the convective nonlinearity dominates and the flow may
become turbulent.

\begin{remark}
  The regularity problem is independent of the Reynolds number in the following sense:
  the Clay Millennium Problem asks whether smooth solutions remain smooth for \emph{all}
  $\nu > 0$, not merely for small Reynolds numbers.  The difficulty lies entirely in the
  three-dimensional geometry of the nonlinear term.
\end{remark}

% ---------------------------------------------------------------------------
\section{Weak formulation}\label{sec:ns-weak}
% ---------------------------------------------------------------------------

The strong formulation~\eqref{eq:ns-momentum}--\eqref{eq:ns-initial} requires $u$ to be
twice differentiable in space and once in time.  To construct solutions with less
regularity, we pass to a weak formulation by testing against smooth, compactly
supported, divergence-free vector fields.

\subsection{Derivation of the weak form}

Let $\varphi \in \Ccinf(\R^{3} \times [0, T); \R^{3})$ with
$\divergence \varphi = 0$.  Multiply the momentum equation~\eqref{eq:ns-momentum} by
$\varphi$, integrate over $\R^{3} \times (0, T)$, and integrate by parts.

The time derivative term gives
\[
  \int_{0}^{T} \int_{\R^{3}} \partial_{t} u \cdot \varphi \,\dd x \,\dd t
  \;=\;
  -\int_{0}^{T} \int_{\R^{3}} u \cdot \partial_{t} \varphi \,\dd x \,\dd t
  \;-\;
  \int_{\R^{3}} u_{0} \cdot \varphi(x, 0) \,\dd x,
\]
using $\varphi(\cdot, T) = 0$ and $u(\cdot, 0) = u_{0}$.

The viscous term, after integration by parts, becomes
\[
  \int_{0}^{T} \int_{\R^{3}} \nu \Delta u \cdot \varphi \,\dd x \,\dd t
  \;=\;
  -\nu \int_{0}^{T} \int_{\R^{3}} \nabla u : \nabla \varphi \,\dd x \,\dd t,
\]
where $\nabla u : \nabla \varphi = \sum_{i,j} \partial_{j} u_{i}\,\partial_{j} \varphi_{i}$.

The pressure term vanishes:
$\int_{\R^{3}} \nabla p \cdot \varphi \,\dd x = -\int_{\R^{3}} p\,\divergence \varphi
\,\dd x = 0$,
since $\divergence \varphi = 0$.  This is the key advantage of testing against
divergence-free functions: the pressure is eliminated entirely.

The convective term, using the conservative form~\eqref{eq:convective-conservative},
integrates by parts to
\[
  \int_{0}^{T} \int_{\R^{3}} (u \cdot \nabla) u \cdot \varphi \,\dd x \,\dd t
  \;=\;
  -\int_{0}^{T} \int_{\R^{3}} (u \otimes u) : \nabla \varphi \,\dd x \,\dd t,
\]
where $(u \otimes u) : \nabla \varphi = \sum_{i,j} u_{i} u_{j}\,\partial_{j} \varphi_{i}$.

Collecting terms yields the weak formulation.

\subsection{Definition of weak solution}

\begin{definition}[Weak solution of Navier--Stokes%
  \footnote{Lean 4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/WeakNSSolution.lean}{\texttt{NavierStokes.IsWeakNSSolution}}}]%
  \label{def:weak-ns}
  Let $u_{0} \in \Lsig(\R^{3})$, $f \in \Lp{2}(0, T; H^{-1}(\R^{3}))$, and
  $0 < T \le \infty$.  A vector field
  \[
    u \in \Lp{\infty}(0, T; \Lp{2}(\R^{3}))
    \cap \Lp{2}(0, T; \Sobz{1}(\R^{3}))
  \]
  with $\divergence u = 0$ in $\mathcal{D}'(\R^{3} \times (0, T))$ is a \emph{weak
  solution} of the Navier--Stokes equations if for every
  $\varphi \in \Ccinf(\R^{3} \times [0, T); \R^{3})$ with $\divergence \varphi = 0$,
  \begin{equation}\label{eq:ns-weak}
    \begin{aligned}
      &-\int_{0}^{T} \int_{\R^{3}} u \cdot \partial_{t} \varphi \,\dd x \,\dd t
      \;-\;
      \int_{\R^{3}} u_{0} \cdot \varphi(x, 0) \,\dd x \\
      &\qquad
      +\;\int_{0}^{T} \int_{\R^{3}} (u \otimes u) : \nabla \varphi \,\dd x \,\dd t
      \;+\;
      \nu \int_{0}^{T} \int_{\R^{3}} \nabla u : \nabla \varphi \,\dd x \,\dd t \\
      &\qquad\qquad
      =\;
      \int_{0}^{T} \ip{f}{\varphi}_{H^{-1} \times \Sobz{1}} \,\dd t.
    \end{aligned}
  \end{equation}
\end{definition}

\begin{remark}[Well-definedness of the nonlinear term]\label{rem:nonlinear-welldefined}
  The nonlinear term $\int (u \otimes u) : \nabla \varphi$ is well-defined when
  $u \in \Lp{2}(0, T; \Sobz{1}(\R^{3}))$.  Indeed, by the Sobolev embedding
  $\Sob{1}(\R^{3}) \hookrightarrow \Lp{6}(\R^{3})$ (\Cref{thm:sobolev-subcritical}
  with $n = 3$, $p = 2$), $u(\cdot, t) \in \Lp{6}(\R^{3})$ for a.e.\ $t$, so
  $u \otimes u \in \Lp{3}(\R^{3})$ and
  $\nabla \varphi \in \Lp{3/2}(\R^{3})$ (since $\varphi$ is smooth and compactly
  supported), whence the integral is finite by H\"{o}lder's inequality.  More precisely,
  $u \otimes u \in \Lp{1}(0, T; \Lp{3}(\R^{3}))$ and the pairing is controlled by the
  energy class norms.
\end{remark}

\begin{remark}[Pressure recovery]
  The pressure $p$ does not appear in the weak formulation~\eqref{eq:ns-weak} because
  we test against divergence-free functions.  Once a weak solution $u$ is obtained, the
  pressure can be recovered (up to an additive function of time) by solving the Poisson
  equation $-\Delta p = \divergence\bigl[(u \cdot \nabla) u\bigr]$ in the sense of
  distributions.  On $\R^{3}$, this is accomplished via the Calder\'{o}n--Zygmund theory
  of singular integrals, yielding $\nabla p \in \Lp{3/2}(\R^{3})$ when $u \in
  \Lp{6}(\R^{3})$.
\end{remark}

% ---------------------------------------------------------------------------
\section{The Galerkin method}\label{sec:galerkin}
% ---------------------------------------------------------------------------

The Galerkin method reduces the Navier--Stokes initial-value problem to a
sequence of finite-dimensional ordinary differential equations whose solutions
approximate the desired weak solution.  On a bounded domain
$\Omega \subset \R^{3}$, one typically projects onto eigenfunctions of the Stokes
operator, exploiting its compact resolvent and discrete spectrum.  On $\R^{3}$ the
Stokes operator has continuous spectrum, so that approach is unavailable.  We
instead construct an approximation basis directly from the separability of
$\Lsig(\R^{3})$, then carry out the Galerkin procedure, derive uniform a~priori
estimates, and pass to the limit using a localisation argument that replaces the
global compactness afforded by Rellich--Kondrachov on bounded domains.

\subsection{Approximation basis on $\R^{3}$}

The space $\Ccinf(\R^{3};\R^{3}) \cap \Lsig(\R^{3})$ is dense in $\Lsig(\R^{3})$
(a consequence of the Helmholtz decomposition, \Cref{thm:helmholtz}).
Since $\Lsig(\R^{3})$ is a separable Hilbert space, this dense subspace
contains a countable dense subset
$\{\varphi_{k}\}_{k=1}^{\infty}$.  Applying the Gram--Schmidt procedure to
$\{\varphi_{k}\}_{k=1}^{\infty}$ in the $\Lp{2}(\R^{3})$ inner product produces an
orthonormal sequence $\{w_{k}\}_{k=1}^{\infty}$ in $\Lsig(\R^{3})$ satisfying:
\begin{enumerate}[label=(\roman*)]
  \item each $w_{k}$ belongs to $\Ccinf(\R^{3};\R^{3})$ and is divergence-free;
  \item $\ip{w_{j}}{w_{k}}_{\Lp{2}} = \delta_{jk}$ for all $j,k \in \N$;
  \item $\overline{\operatorname{span}\{w_{k} : k \in \N\}} = \Lsig(\R^{3})$.
\end{enumerate}
Property~(i) holds because each $w_{k}$ is a finite linear combination of the
$\varphi_{j}$, all of which are smooth, compactly supported, and divergence-free;
the linear combination inherits each property.  Property~(iii) follows because
$\operatorname{span}\{w_{k}\}$ contains $\operatorname{span}\{\varphi_{k}\}$
(each $\varphi_{k}$ can be recovered from $w_{1}, \ldots, w_{k}$ by inverting the
Gram--Schmidt triangular system), and the latter is dense in $\Lsig(\R^{3})$ by
construction.

\begin{remark}[Bounded-domain alternative]
  On a bounded domain $\Omega$ with smooth boundary, the Stokes operator
  $A = -\Leray\Delta$ has compact
  resolvent\footnote{Lean~4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/WeakNSSolution.lean}{\texttt{NavierStokes.StokesOperator}}.}
  and therefore possesses a complete orthonormal system of eigenfunctions in
  $\Lsig(\Omega)$ with eigenvalues
  $0 < \lambda_{1} \le \lambda_{2} \le \cdots \to \infty$.  This eigenbasis is the
  standard choice for Galerkin approximation in the bounded setting; see
  Temam~\cite{temam2001}, Chapter~III.  On $\R^{3}$ the Stokes operator is still
  well-defined as a self-adjoint operator on $\Lsig(\R^{3})$, but its spectrum is
  purely continuous, so eigenfunctions in the classical sense do not exist.  The
  Gram--Schmidt construction above circumvents this difficulty entirely.
\end{remark}

\subsection{Finite-dimensional approximation}

For each $m \in \N$, define the Galerkin subspace
\begin{equation}\label{eq:galerkin-subspace}
  V_{m}
  \;=\;
  \operatorname{span}\{w_{1}, \ldots, w_{m}\}
  \;\subset\;
  \Lsig(\R^{3}),
\end{equation}
and let $P_{m} \colon \Lsig(\R^{3}) \to V_{m}$ denote the orthogonal projection.
We seek an approximate solution of the form
\begin{equation}\label{eq:galerkin-ansatz}
  u_{m}(t)
  \;=\;
  \sum_{k=1}^{m} c_{k}^{(m)}(t)\,w_{k},
\end{equation}
where the coefficient functions $c_{k}^{(m)} \colon [0, T_{m}) \to \R$ are to be
determined.  The trilinear form%
  \footnote{Lean 4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/TrilinearForm.lean}{\texttt{NavierStokes.trilinearForm}}}
\begin{equation}\label{eq:trilinear-form}
  b(u, v, w)
  \;=\;
  \int_{\R^{3}} (u \cdot \nabla) v \cdot w \,\dd x
\end{equation}
is well-defined whenever the integrand is summable.  We require $u_{m}$ to satisfy
the projected equations: for each $k = 1, \ldots, m$,
\begin{align}\label{eq:galerkin-ode}
  \ip{\partial_{t} u_{m}}{w_{k}}_{\Lp{2}}
  &\;+\;
  \nu\,\ip{\nabla u_{m}}{\nabla w_{k}}_{\Lp{2}}
  \;+\;
  b(u_{m}, u_{m}, w_{k})
  \;=\;
  \ip{f}{w_{k}}_{\Lp{2}}, \notag \\
  u_{m}(0) &\;=\; P_{m} u_{0}.
\end{align}

The antisymmetry of the trilinear form is the key structural property underpinning
all energy estimates.

\begin{lemma}[Antisymmetry of the trilinear form%
  \footnote{Lean 4 (machine-checked): \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/TrilinearForm.lean}{\texttt{NavierStokes.trilinearForm\_antisymmetric}}}]%
  \label{lem:trilinear-antisymmetry}
  Let $u \in \Ccinf(\R^{3}; \R^{3})$ with $\divergence u = 0$, and let
  $v \in \Sob{1}(\R^{3}; \R^{3})$.  Then
  \begin{equation}\label{eq:trilinear-antisymmetry}
    b(u, v, v) \;=\; 0.
  \end{equation}
  More generally, $b(u, v, w) = -b(u, w, v)$ for all
  $v, w \in \Sob{1}(\R^{3}; \R^{3})$.
\end{lemma}

\begin{proof}
  Write out the trilinear form componentwise:
  \[
    b(u, v, v)
    \;=\;
    \sum_{i,j=1}^{3} \int_{\R^{3}} u_{j}\,\partial_{j} v_{i}\,v_{i} \,\dd x
    \;=\;
    \frac{1}{2} \sum_{j=1}^{3} \int_{\R^{3}} u_{j}\,\partial_{j}\abs{v}^{2} \,\dd x.
  \]
  Since $u$ is smooth and compactly supported, integration by parts yields
  \[
    \frac{1}{2} \sum_{j=1}^{3} \int_{\R^{3}} u_{j}\,\partial_{j}\abs{v}^{2} \,\dd x
    \;=\;
    -\frac{1}{2} \int_{\R^{3}} (\divergence u)\,\abs{v}^{2} \,\dd x
    \;=\;
    0,
  \]
  where the boundary terms vanish because $u$ is compactly supported and
  $v \in \Sob{1}(\R^{3}; \R^{3})$, and the final equality uses
  $\divergence u = 0$.  The general antisymmetry identity $b(u, v, w) = -b(u, w, v)$
  follows by polarisation: apply the vanishing result to $v + w$ and $v - w$,
  then subtract.
\end{proof}

\begin{remark}[Lean formalisation]\label{rem:trilinear-lean}
  The machine-checked proof establishes a dual formulation: for
  \emph{distributionally} divergence-free $u$ (no pointwise regularity required)
  and $v \in \Ccinf(\R^{3}; \R^{3})$, one has $b(u, v, v) = 0$.  The proof
  applies the distributional divergence-free condition
  $\sum_{j} \int u_{j}\,\partial_{j}\varphi = 0$ separately to each test function
  $\varphi_{k} = v_{k}^{2}$, bypassing integration by parts entirely.
  Since the Galerkin basis functions $\{w_{k}\}$ lie in
  $\Ccinf(\R^{3}) \cap \Lsig(\R^{3})$, both formulations apply to the energy
  estimate $b(u_{m}, u_{m}, u_{m}) = 0$.  The extension of
  Lemma~\ref{lem:trilinear-antisymmetry} from $\Ccinf$ to $\Sob{1}$ is by
  density and continuity of $b$ in the $\Sob{1}$ topology (not yet formalised).
\end{remark}

Substituting the ansatz~\eqref{eq:galerkin-ansatz} into~\eqref{eq:galerkin-ode}
and using the orthonormality $\ip{w_{j}}{w_{k}}_{\Lp{2}} = \delta_{jk}$ yields an
ODE system for the coefficient vector
\[
  c^{(m)}(t)
  \;=\;
  \bigl(c_{1}^{(m)}(t), \ldots, c_{m}^{(m)}(t)\bigr)^{\top}
  \in \R^{m}\colon
\]
\[
  \dot{c}_{k}^{(m)}
  \;=\;
  -\nu \sum_{j=1}^{m} A_{kj}\,c_{j}^{(m)}
  \;-\;
  \sum_{i,j=1}^{m} B_{kij}\,c_{i}^{(m)}\,c_{j}^{(m)}
  \;+\;
  f_{k}(t),
  \quad
  k = 1, \ldots, m,
\]
where $A_{kj} = \ip{\nabla w_{j}}{\nabla w_{k}}_{\Lp{2}}$,
$B_{kij} = b(w_{i}, w_{j}, w_{k})$, and
$f_{k}(t) = \ip{f(t)}{w_{k}}_{\Lp{2}}$.  The right-hand side is locally Lipschitz
in $c^{(m)}$ (it is quadratic) and measurable in $t$, so the Cauchy--Lipschitz
theorem applies.

\begin{proposition}[Local existence of Galerkin approximations]%
  \label{prop:galerkin-local}
  For each $m \in \N$, there exists $T_{m} > 0$ and a unique absolutely continuous
  function $c^{(m)} \colon [0, T_{m}) \to \R^{m}$ satisfying the ODE system above with
  initial data $c_{k}^{(m)}(0) = \ip{u_{0}}{w_{k}}_{\Lp{2}}$.  The corresponding
  $u_{m} \in C([0, T_{m}); V_{m})$ satisfies~\eqref{eq:galerkin-ode} for a.e.\
  $t \in (0, T_{m})$.
\end{proposition}

\begin{proof}
  The coefficients $A_{kj}$ are constants (each $w_{k}$ is smooth and compactly
  supported, so $\ip{\nabla w_{j}}{\nabla w_{k}}_{\Lp{2}} < \infty$).  Likewise, the
  $B_{kij}$ are finite constants because
  $w_{i}, w_{j}, w_{k} \in \Ccinf(\R^{3}; \R^{3})$.
  The forcing terms $f_{k}(t) = \ip{f(t)}{w_{k}}_{\Lp{2}}$ are in
  $\Lp{2}(0, T)$ whenever $f \in \Lp{2}(0, T; \Sob{-1}(\R^{3}))$, and in
  particular are locally integrable.  The right-hand side of the ODE is therefore
  a locally Lipschitz function of $c^{(m)}$ plus a locally integrable function of
  $t$, so the Cauchy--Lipschitz (Picard--Lindel\"{o}f) theorem gives a unique
  maximal solution on some interval $[0, T_{m})$.
\end{proof}

\subsection{A priori estimates}\label{subsec:a-priori}

The following energy estimate is the heart of the Galerkin method.  It provides
bounds that are uniform in $m$ and sufficient to extract a convergent subsequence.
For simplicity of exposition, we set $f = 0$; the case of nonzero forcing requires
only minor modifications (absorbing $f$ via the Cauchy--Schwarz and Young
inequalities).

\begin{theorem}[Energy estimate for Galerkin approximations]%
  \label{thm:galerkin-energy}
  For $f = 0$, the Galerkin approximation $u_{m}$ satisfies
  \begin{equation}\label{eq:galerkin-energy-estimate}
    \frac{1}{2}\norm{u_{m}(t)}_{\Lp{2}}^{2}
    \;+\;
    \nu \int_{0}^{t} \norm{\nabla u_{m}(s)}_{\Lp{2}}^{2} \,\dd s
    \;=\;
    \frac{1}{2}\norm{P_{m} u_{0}}_{\Lp{2}}^{2}
    \;\le\;
    \frac{1}{2}\norm{u_{0}}_{\Lp{2}}^{2}
  \end{equation}
  for all $t \in [0, T_{m})$.
\end{theorem}

\begin{proof}
  Multiply the Galerkin equation~\eqref{eq:galerkin-ode} by $c_{k}^{(m)}(t)$ and sum
  over $k = 1, \ldots, m$.  Since $u_{m} = \sum_{k} c_{k}^{(m)} w_{k}$, the left-hand
  side becomes
  $\ip{\partial_{t} u_{m}}{u_{m}}_{\Lp{2}} + \nu\,\norm{\nabla u_{m}}_{\Lp{2}}^{2}
  + b(u_{m}, u_{m}, u_{m})$.  By \Cref{lem:trilinear-antisymmetry},
  $b(u_{m}, u_{m}, u_{m}) = 0$.  The first term equals
  $\frac{1}{2}\frac{\dd}{\dd t}\norm{u_{m}}_{\Lp{2}}^{2}$.  With $f = 0$, the
  right-hand side vanishes, giving
  \[
    \frac{1}{2}\frac{\dd}{\dd t}\norm{u_{m}(t)}_{\Lp{2}}^{2}
    \;+\;
    \nu\,\norm{\nabla u_{m}(t)}_{\Lp{2}}^{2}
    \;=\;
    0.
  \]
  Integrating from $0$ to $t$ and using $u_{m}(0) = P_{m} u_{0}$ yields
  \eqref{eq:galerkin-energy-estimate}.  The inequality
  $\norm{P_{m} u_{0}}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$ follows from the fact that
  $P_{m}$ is an orthogonal projection.
\end{proof}

\begin{corollary}[Global existence and uniform bounds]\label{cor:galerkin-global}
  The Galerkin approximation $u_{m}$ extends to all of $[0, T]$ for every $T > 0$, and
  the following uniform bounds hold (independent of $m$):
  \begin{equation}\label{eq:galerkin-uniform-bounds}
    \sup_{t \in [0, T]} \norm{u_{m}(t)}_{\Lp{2}}^{2}
    \;\le\;
    \norm{u_{0}}_{\Lp{2}}^{2},
    \qquad
    \int_{0}^{T} \norm{\nabla u_{m}(t)}_{\Lp{2}}^{2} \,\dd t
    \;\le\;
    \frac{1}{2\nu}\norm{u_{0}}_{\Lp{2}}^{2}.
  \end{equation}
\end{corollary}

\begin{proof}
  The energy estimate~\eqref{eq:galerkin-energy-estimate} shows that
  $\norm{u_{m}(t)}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$ for all $t \in [0, T_{m})$.
  In terms of the coefficient vector $c^{(m)}(t)$, this reads
  $\abs{c^{(m)}(t)} = \norm{u_{m}(t)}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$, so
  $c^{(m)}(t)$ remains bounded.  By the continuation criterion for ODE solutions (a
  solution that remains bounded extends beyond its maximal interval),
  $T_{m} = +\infty$.  The bound on $\int \norm{\nabla u_{m}}^{2}$ follows directly
  from~\eqref{eq:galerkin-energy-estimate} by dropping the non-negative first term.
\end{proof}

We also need a bound on the time derivatives to apply compactness arguments.

\begin{lemma}[Time derivative bound]\label{lem:time-derivative-bound}
  The sequence $\{\partial_{t} u_{m}\}_{m=1}^{\infty}$ satisfies
  \begin{equation}\label{eq:time-derivative-bound}
    \sup_{m \in \N}
    \norm{\partial_{t} u_{m}}_{\Lp{4/3}(0,T;\,\Sob{-1}(\R^{3}))}
    \;\le\;
    C\bigl(\nu, \norm{u_{0}}_{\Lp{2}}\bigr).
  \end{equation}
\end{lemma}

\begin{proof}
  From~\eqref{eq:galerkin-ode}, for any $\phi \in \Sobz{1}(\R^{3}; \R^{3})$ with
  $\divergence \phi = 0$,
  \[
    \ip{\partial_{t} u_{m}}{\phi}_{\Sob{-1}, \Sobz{1}}
    \;=\;
    -\nu\,\ip{\nabla u_{m}}{\nabla \phi}_{\Lp{2}}
    \;-\;
    b(u_{m}, u_{m}, \phi).
  \]
  The viscous term satisfies
  $\abs{\nu\,\ip{\nabla u_{m}}{\nabla \phi}_{\Lp{2}}} \le
  \nu\,\norm{\nabla u_{m}}_{\Lp{2}}\,\norm{\nabla \phi}_{\Lp{2}}$.
  For the nonlinear term, H\"{o}lder's inequality with the Sobolev embedding
  $\Sobz{1}(\R^{3}) \hookrightarrow \Lp{6}(\R^{3})$
  (\Cref{thm:sobolev-subcritical}) and interpolation give
  \[
    \abs{b(u_{m}, u_{m}, \phi)}
    \;\le\;
    C\,\norm{u_{m}}_{\Lp{2}}^{1/2}\,\norm{\nabla u_{m}}_{\Lp{2}}^{3/2}
    \,\norm{\nabla \phi}_{\Lp{2}}.
  \]
  Dividing by $\norm{\nabla \phi}_{\Lp{2}}$ and taking the supremum over $\phi$ yields
  \begin{align*}
    \norm{\partial_{t} u_{m}(t)}_{\Sob{-1}}
    &\;\le\;
    \nu\,\norm{\nabla u_{m}(t)}_{\Lp{2}} \\
    &\qquad
    +\;C\,\norm{u_{0}}_{\Lp{2}}^{1/2}\,
    \norm{\nabla u_{m}(t)}_{\Lp{2}}^{3/2},
  \end{align*}
  using $\norm{u_{m}}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$ from
  \Cref{cor:galerkin-global}.  Raising to the power $4/3$ and integrating in time,
  the dominant term $\norm{\nabla u_{m}}^{3/2}$ raised to the power $4/3$
  gives $\norm{\nabla u_{m}}^{2}$, so
  \[
    \int_{0}^{T} \norm{\partial_{t} u_{m}(t)}_{\Sob{-1}}^{4/3} \,\dd t
    \;\le\;
    C\Bigl(\nu, \norm{u_{0}}_{\Lp{2}},
    \int_{0}^{T} \norm{\nabla u_{m}}_{\Lp{2}}^{2} \,\dd t\Bigr),
  \]
  and the right-hand side is bounded uniformly in $m$
  by~\eqref{eq:galerkin-uniform-bounds}.
\end{proof}

\subsection{Passage to the limit}

We now extract a subsequence of Galerkin approximations that converges to a
weak solution of the Navier--Stokes equations on $\R^{3} \times (0, T)$.  The
principal difficulty, compared with the bounded-domain case, is that the
embedding $\Sobz{1}(\R^{3}) \hookrightarrow \Lp{2}(\R^{3})$ is continuous but not
compact.  We overcome this by a localisation and diagonal argument.

\begin{theorem}[Aubin--Lions, local version]\label{thm:aubin-lions}
  Let $B_{R} = \{x \in \R^{3} : \abs{x} < R\}$ for $R > 0$.  Suppose
  $\{v_{m}\}_{m=1}^{\infty}$ is uniformly bounded in
  $\Lp{2}(0, T; \Sobz{1}(\R^{3}))$ with $\{\partial_{t} v_{m}\}$ uniformly bounded in
  $\Lp{4/3}(0, T; \Sob{-1}(\R^{3}))$.  Then for each fixed $R > 0$, the restrictions
  $\{v_{m}|_{B_{R}}\}$ are relatively compact in $\Lp{2}(0, T; \Lp{2}(B_{R}))$.
\end{theorem}

\begin{proof}
  Fix $R > 0$.  The restriction map $\Sobz{1}(\R^{3}) \to \Sob{1}(B_{R})$ is
  continuous and the embedding $\Sob{1}(B_{R}) \hookrightarrow \Lp{2}(B_{R})$ is
  compact by the Rellich--Kondrachov theorem (\Cref{thm:rellich-kondrachov}), since
  $B_{R}$ is a bounded Lipschitz domain.  Likewise,
  $\Lp{2}(B_{R}) \hookrightarrow \Sob{-1}(B_{R})$ continuously.  The restrictions
  $v_{m}|_{B_{R}}$ are therefore uniformly bounded in
  $\Lp{2}(0, T; \Sob{1}(B_{R}))$, and $\partial_{t}(v_{m}|_{B_{R}})$ are uniformly
  bounded in $\Lp{4/3}(0, T; \Sob{-1}(B_{R}))$ (since restriction commutes with
  time differentiation in the distributional sense, and the restriction of an
  $\Sob{-1}(\R^{3})$ distribution to $B_{R}$ has no larger $\Sob{-1}(B_{R})$ norm).
  The classical Aubin--Lions lemma (with the triple
  $\Sob{1}(B_{R}) \Subset \Lp{2}(B_{R}) \hookrightarrow \Sob{-1}(B_{R})$
  and exponents $p = 2$, $q = 4/3$) yields relative
  compactness of $\{v_{m}|_{B_{R}}\}$ in $\Lp{2}(0, T; \Lp{2}(B_{R}))$.
\end{proof}

\begin{theorem}[Convergence of Galerkin approximations]\label{thm:galerkin-convergence}
  There exist a subsequence $\{u_{m_{j}}\}_{j=1}^{\infty}$ and a function
  \[
    u \in \Lp{\infty}(0, T; \Lsig(\R^{3}))
    \cap \Lp{2}(0, T; \Sobz{1}(\R^{3}; \R^{3}))
  \]
  such that:
  \begin{enumerate}[label=(\roman*)]
    \item $u_{m_{j}} \rightharpoonup u$ weakly in
          $\Lp{2}(0, T; \Sobz{1}(\R^{3}))$;
    \item $u_{m_{j}} \overset{*}{\rightharpoonup} u$ weak-$*$ in
          $\Lp{\infty}(0, T; \Lp{2}(\R^{3}))$;
    \item $u_{m_{j}} \to u$ strongly in
          $\Lp{2}(0, T; \Lp{2}(K))$ for every compact set $K \subset \R^{3}$.
  \end{enumerate}
\end{theorem}

\begin{proof}
  The uniform bounds~\eqref{eq:galerkin-uniform-bounds} show that
  $\{u_{m}\} \subset \Lp{\infty}(0, T; \Lsig(\R^{3})) \cap
  \Lp{2}(0, T; \Sobz{1}(\R^{3}))$, both bounded.  By the Banach--Alaoglu theorem,
  there exist a subsequence (still denoted $\{u_{m}\}$) and $u$ in the indicated
  spaces satisfying (i) and~(ii).

  For (iii), we use a diagonal argument.  By \Cref{thm:aubin-lions} applied with
  $R = 1$, there is a subsequence $\{u_{m}^{(1)}\}$ such that
  $u_{m}^{(1)}|_{B_{1}} \to u|_{B_{1}}$ strongly in
  $\Lp{2}(0, T; \Lp{2}(B_{1}))$.
  Applying \Cref{thm:aubin-lions} again with $R = 2$ to the subsequence
  $\{u_{m}^{(1)}\}$, we extract a further subsequence $\{u_{m}^{(2)}\}$ converging
  strongly on $B_{2}$.  Continuing inductively, for each $n \in \N$ we obtain a
  subsequence $\{u_{m}^{(n)}\}$ of $\{u_{m}^{(n-1)}\}$ converging strongly in
  $\Lp{2}(0, T; \Lp{2}(B_{n}))$.  The diagonal subsequence
  $u_{m_{j}} = u_{j}^{(j)}$ then converges strongly in
  $\Lp{2}(0, T; \Lp{2}(B_{n}))$ for every $n \in \N$.  Since every compact
  $K \subset \R^{3}$ is contained in some $B_{n}$, conclusion~(iii) follows.
\end{proof}

It remains to verify that the limit $u$ satisfies the weak formulation of the
Navier--Stokes equations.

\begin{theorem}[Passage to the limit]\label{thm:passage-limit}
  The function $u$ obtained in \Cref{thm:galerkin-convergence} is a weak solution
  of the Navier--Stokes equations on $\R^{3} \times (0, T)$ in the sense
  of~\eqref{eq:ns-weak} (with $f = 0$).
\end{theorem}

\begin{proof}
  Fix $\phi \in \Ccinf(\R^{3} \times [0, T); \R^{3})$ with
  $\divergence \phi = 0$.  Choose $R > 0$ large enough that
  $\supp \phi \subset B_{R} \times [0, T)$.  For each $m$, let
  $\phi_{m}(t) = \sum_{k=1}^{m} \ip{\phi(t)}{w_{k}} w_{k}$ be the projection of
  $\phi$ onto $V_{m}$.  Since $\{w_{k}\}$ is complete in $\Lsig(\R^{3})$,
  $\phi_{m} \to \phi$ in every relevant norm.

  The Galerkin equation~\eqref{eq:galerkin-ode} with test function $\phi_{m}$, after
  integration by parts in time, gives
  \begin{align*}
    &-\int_{0}^{T} \ip{u_{m}}{\partial_{t} \phi_{m}}_{\Lp{2}} \,\dd t
    \;-\;
    \ip{P_{m} u_{0}}{\phi_{m}(0)}_{\Lp{2}} \\
    &\qquad
    +\;\int_{0}^{T} b(u_{m}, u_{m}, \phi_{m}) \,\dd t
    \;+\;
    \nu \int_{0}^{T} \ip{\nabla u_{m}}{\nabla \phi_{m}}_{\Lp{2}} \,\dd t
    \;=\;
    0.
  \end{align*}
  We pass each term to the limit along the subsequence $\{u_{m_{j}}\}$.

  The linear terms converge by weak and weak-$*$ convergence.  For the
  time-derivative term, $u_{m} \rightharpoonup u$ in $\Lp{2}(0, T; \Lp{2})$ and
  $\partial_{t}\phi_{m} \to \partial_{t}\phi$ strongly, so
  \[
    \int_{0}^{T} \ip{u_{m}}{\partial_{t}\phi_{m}} \,\dd t
    \;\to\;
    \int_{0}^{T} \ip{u}{\partial_{t}\phi} \,\dd t.
  \]
  Similarly, weak convergence of $\nabla u_{m}$ in $\Lp{2}(0, T; \Lp{2})$ gives
  \[
    \int_{0}^{T} \ip{\nabla u_{m}}{\nabla \phi_{m}} \,\dd t
    \;\to\;
    \int_{0}^{T} \ip{\nabla u}{\nabla \phi} \,\dd t.
  \]
  The initial data term converges because $P_{m} u_{0} \to u_{0}$ in $\Lp{2}$
  and $\phi_{m}(0) \to \phi(0)$.

  The nonlinear term is the critical passage.  Since
  $\supp \phi \subset B_{R} \times [0, T)$,
  \[
    \int_{0}^{T} b(u_{m}, u_{m}, \phi_{m}) \,\dd t
    \;=\;
    -\int_{0}^{T}\!\!\int_{B_{R}}
    (u_{m} \otimes u_{m}) : \nabla \phi_{m} \,\dd x \,\dd t.
  \]
  We decompose
  $u_{m} \otimes u_{m} - u \otimes u = (u_{m} - u) \otimes u_{m} +
  u \otimes (u_{m} - u)$.
  The strong convergence $u_{m_{j}} \to u$ in
  $\Lp{2}(0, T; \Lp{2}(B_{R}))$ from
  \Cref{thm:galerkin-convergence}(iii), combined with the uniform bound
  $\norm{u_{m}}_{\Lp{\infty}(0, T; \Lp{2})} \le \norm{u_{0}}_{\Lp{2}}$, gives
  $(u_{m_{j}} - u) \otimes u_{m_{j}} \to 0$ in
  $\Lp{1}(0, T; \Lp{1}(B_{R}))$.  For the second factor,
  $u \otimes (u_{m_{j}} - u) \to 0$ by the weak convergence
  $u_{m_{j}} \rightharpoonup u$ in $\Lp{2}(0, T; \Sobz{1})$, since
  $(u \cdot \nabla)\phi$ is a fixed bounded function supported in
  $B_{R} \times [0, T)$.  Since
  $\nabla \phi_{m} \to \nabla \phi$ in
  $\Lp{\infty}(0, T; \Lp{\infty}(\R^{3}))$, we conclude
  \[
    \int_{0}^{T} b(u_{m_{j}}, u_{m_{j}}, \phi_{m_{j}}) \,\dd t
    \;\to\;
    \int_{0}^{T} b(u, u, \phi) \,\dd t.
  \]
  Combining all terms, $u$ satisfies~\eqref{eq:ns-weak}.
\end{proof}

% ---------------------------------------------------------------------------
\section{Energy inequality}\label{sec:energy-inequality}
% ---------------------------------------------------------------------------

The Galerkin approximations satisfy an energy \emph{equality}
(\Cref{thm:galerkin-energy}), but after passage to the limit, only an
\emph{inequality} survives.

\subsection{Statement and derivation}

\begin{theorem}[Energy inequality%
  \footnote{Lean 4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/EnergyInequality.lean}{\texttt{NavierStokes.SatisfiesEnergyInequality}}}]%
  \label{thm:energy-inequality}
  Let $u$ be the weak solution obtained in \Cref{thm:passage-limit}.  Then for
  almost every $t \in (0, T)$,
  \begin{equation}\label{eq:energy-inequality}
    \frac{1}{2}\norm{u(t)}_{\Lp{2}}^{2}
    \;+\;
    \nu \int_{0}^{t} \norm{\nabla u(s)}_{\Lp{2}}^{2} \,\dd s
    \;\le\;
    \frac{1}{2}\norm{u_{0}}_{\Lp{2}}^{2}.
  \end{equation}
\end{theorem}

\begin{proof}
  The Galerkin approximations satisfy the energy equality
  \[
    \frac{1}{2}\norm{u_{m}(t)}_{\Lp{2}}^{2}
    \;+\;
    \nu \int_{0}^{t} \norm{\nabla u_{m}(s)}_{\Lp{2}}^{2} \,\dd s
    \;=\;
    \frac{1}{2}\norm{P_{m} u_{0}}_{\Lp{2}}^{2}
  \]
  for all $t \in [0, T]$, by \Cref{thm:galerkin-energy}.  We pass each side to the
  limit separately.

  The right-hand side converges: $\norm{P_{m} u_{0}}_{\Lp{2}} \to
  \norm{u_{0}}_{\Lp{2}}$ as $m \to \infty$, since $P_{m} u_{0} \to u_{0}$ in
  $\Lp{2}(\Omega)$ (completeness of the eigenfunctions $\{w_{k}\}$).

  For the left-hand side, the strong convergence $u_{m} \to u$ in
  $\Lp{2}(0,T;\Lp{2})$ from \Cref{thm:galerkin-convergence}(iii) gives,
  after passing to a subsequence, $u_{m}(t) \to u(t)$ in $\Lp{2}$ for
  a.e.\ $t$.  For such $t$,
  \[
    \norm{u(t)}_{\Lp{2}}
    \;=\;
    \lim_{m \to \infty} \norm{u_{m}(t)}_{\Lp{2}}.
  \]
  For the dissipation term, weak convergence
  $\nabla u_{m} \rightharpoonup \nabla u$ in $\Lp{2}(0, T; \Lp{2})$ and
  weak lower semicontinuity of the norm yield
  \[
    \int_{0}^{t} \norm{\nabla u(s)}_{\Lp{2}}^{2} \,\dd s
    \;\le\;
    \liminf_{m \to \infty}
    \int_{0}^{t} \norm{\nabla u_{m}(s)}_{\Lp{2}}^{2} \,\dd s.
  \]
  Combining, for a.e.\ $t$,
  \begin{align*}
    \frac{1}{2}\norm{u(t)}_{\Lp{2}}^{2}
    &+ \nu \int_{0}^{t}\norm{\nabla u(s)}_{\Lp{2}}^{2}\,\dd s \\
    &\;\le\;
    \liminf_{m \to \infty}\Bigl[
    \frac{1}{2}\norm{u_{m}(t)}_{\Lp{2}}^{2}
    + \nu \int_{0}^{t}\norm{\nabla u_{m}}_{\Lp{2}}^{2}\,\dd s\Bigr] \\
    &\;=\;
    \frac{1}{2}\norm{u_{0}}_{\Lp{2}}^{2}.
  \end{align*}
  This establishes~\eqref{eq:energy-inequality}.
\end{proof}

\subsection{Why equality may fail}

\begin{remark}[Energy equality versus inequality]\label{rem:energy-equality-failure}
  The passage from equality (for Galerkin approximations) to inequality (for the limit)
  arises because weak and weak-$*$ limits do not preserve norms in general: one has only
  the lower semicontinuity $\norm{u} \le \liminf \norm{u_{m}}$, not equality.
  Equality would hold if the convergence were \emph{strong} in $\Lp{\infty}(0, T;
  \Lp{2})$, but the Aubin--Lions lemma provides strong convergence only in
  $\Lp{2}(0, T; \Lp{2})$ (i.e., integrated in time, not pointwise in time).

  Physically, the energy inequality states that kinetic energy can only decrease (or be
  dissipated by viscosity), never spontaneously increase.  The strict inequality would
  correspond to \emph{anomalous dissipation}, a phenomenon related to turbulence and the
  Onsager conjecture.  Whether every Leray--Hopf weak solution satisfies energy
  \emph{equality} remains open in three dimensions.  In two dimensions, energy equality
  is known to hold because the solutions are unique and regular.
\end{remark}

\subsection{The energy inequality with general initial times}

\begin{remark}[Strong energy inequality]\label{rem:strong-energy-inequality}
  The \emph{strong energy inequality} asserts that for almost every $s \ge 0$ and all
  $t \ge s$,
  \begin{equation}\label{eq:strong-energy-inequality}
    \frac{1}{2}\norm{u(t)}_{\Lp{2}}^{2}
    \;+\;
    \nu \int_{s}^{t} \norm{\nabla u(\tau)}_{\Lp{2}}^{2} \,\dd \tau
    \;\le\;
    \frac{1}{2}\norm{u(s)}_{\Lp{2}}^{2}.
  \end{equation}
  This is a stronger statement than~\eqref{eq:energy-inequality} because it holds with
  arbitrary initial time $s$ (not just $s = 0$).  The Leray--Hopf construction in fact
  produces solutions satisfying~\eqref{eq:strong-energy-inequality}; the proof uses the
  same weak lower semicontinuity argument applied to the Galerkin energy equality
  restarted at time $s$.
\end{remark}

% ---------------------------------------------------------------------------
\section{Leray--Hopf weak solutions}\label{sec:leray-hopf-def}
% ---------------------------------------------------------------------------

We now collect the properties established in the preceding sections into a single
definition.

\subsection{The definition}

\begin{definition}[Leray--Hopf weak solution%
  \footnote{Lean 4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/Existence.lean}{\texttt{NavierStokes.IsLerayHopfSolution}}}]%
  \label{def:leray-hopf}
  Let $u_{0} \in \Lsig(\R^{3})$ and $0 < T \le \infty$.  A vector field
  $u : \R^{3} \times (0, T) \to \R^{3}$ is a \emph{Leray--Hopf weak solution}
  of the Navier--Stokes equations with initial data $u_{0}$ if:
  \begin{enumerate}[label=(\roman*)]
    \item \emph{Regularity class.} $u \in \Lp{\infty}(0, T; \Lsig(\R^{3}))
      \cap \Lp{2}(0, T; \Sobz{1}(\R^{3}; \R^{3}))$.
    \item \emph{Weak formulation.} For every
      $\varphi \in \Ccinf(\R^{3} \times [0, T); \R^{3})$ with
      $\divergence \varphi = 0$, the identity~\eqref{eq:ns-weak} holds (with $f = 0$).
    \item \emph{Energy inequality.} For almost every $t \in (0, T)$,
      \[
        \frac{1}{2}\norm{u(t)}_{\Lp{2}}^{2}
        \;+\;
        \nu \int_{0}^{t} \norm{\nabla u(s)}_{\Lp{2}}^{2} \,\dd s
        \;\le\;
        \frac{1}{2}\norm{u_{0}}_{\Lp{2}}^{2}.
      \]
    \item \emph{Weak continuity at $t = 0$.}
      $u(t) \rightharpoonup u_{0}$ in $\Lp{2}(\R^{3})$ as $t \to 0^{+}$.
  \end{enumerate}
\end{definition}

\begin{remark}[Weak versus strong continuity in time]
  Condition (iv) requires only that $u(t)$ converges \emph{weakly} to $u_{0}$ as
  $t \to 0^{+}$.  In fact, the energy inequality (iii) upgrades this to strong
  continuity: the weak convergence $u(t) \rightharpoonup u_{0}$ implies
  $\norm{u_{0}}_{\Lp{2}} \le \liminf_{t \to 0^{+}} \norm{u(t)}_{\Lp{2}}$, while the
  energy inequality gives
  $\limsup_{t \to 0^{+}} \norm{u(t)}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$.  Together,
  $\norm{u(t)}_{\Lp{2}} \to \norm{u_{0}}_{\Lp{2}}$, and in a Hilbert space, weak
  convergence together with convergence of norms implies strong convergence.  Thus
  $u(t) \to u_{0}$ strongly in $\Lp{2}(\R^{3})$ as $t \to 0^{+}$.
\end{remark}

\begin{remark}[Time regularity]\label{rem:time-regularity}
  A Leray--Hopf weak solution $u$ belongs to $C_{w}([0, T); \Lp{2}(\R^{3}))$, the
  space of weakly continuous functions from $[0, T)$ to $\Lp{2}(\R^{3})$.  This means
  that for every $\psi \in \Lp{2}(\R^{3})$, the map $t \mapsto \ip{u(t)}{\psi}_{\Lp{2}}$
  is continuous on $[0, T)$.  The weak continuity follows from the weak
  formulation~\eqref{eq:ns-weak} by choosing test functions of the form
  $\varphi(x, t) = \eta(t)\,\psi(x)$ with $\eta \in \Ccinf([0, T))$.
\end{remark}

\subsection{The existence theorem}

\begin{theorem}[Existence of Leray--Hopf weak solutions%
  \footnote{Lean 4: \href{https://github.com/alejandro-soto-franco/navier-stokes/blob/master/lean/NavierStokes/LerayHopf/Existence.lean}{\texttt{NavierStokes.lerayHopf\_existence}}}]%
  \label{thm:leray-hopf-existence}
  For every $u_{0} \in \Lsig(\R^{3})$ and every $T > 0$ (including $T = \infty$), there
  exists a Leray--Hopf weak solution $u$ of the Navier--Stokes equations on
  $\R^{3} \times (0, T)$ with initial data $u_{0}$, in the sense of
  \Cref{def:leray-hopf}.
\end{theorem}

\begin{proof}
  The proof assembles the results of the preceding sections.  We give the construction
  on a bounded domain $\Omega$, then extend to $\R^{3}$.

  The Galerkin approximations $u_{m}$ are defined by~\eqref{eq:galerkin-ansatz}
  and~\eqref{eq:galerkin-ode}, using the orthonormal basis of $\Lsig(\R^{3})$
  constructed via Gram--Schmidt in \Cref{sec:galerkin}.  By
  \Cref{prop:galerkin-local}, each $u_{m}$ exists
  locally, and by \Cref{cor:galerkin-global}, each $u_{m}$ extends to $[0, T]$ with the
  uniform bounds~\eqref{eq:galerkin-uniform-bounds}.  By
  \Cref{thm:galerkin-convergence}, a subsequence converges to a limit $u$ in the
  senses (i)--(iii) of that theorem.  By \Cref{thm:passage-limit}, $u$ satisfies the weak
  formulation.  By \Cref{thm:energy-inequality}, $u$ satisfies the energy inequality.
  This verifies conditions (i)--(iii) of \Cref{def:leray-hopf}.

  For condition (iv), weak continuity at $t = 0$: the energy inequality gives
  $\limsup_{t \to 0^{+}} \norm{u(t)}_{\Lp{2}} \le \norm{u_{0}}_{\Lp{2}}$, so $(u(t))$
  is bounded in $\Lp{2}$ as $t \to 0^{+}$.  By the Banach--Alaoglu theorem, every
  sequence $t_{k} \to 0^{+}$ has a subsequence along which $u(t_{k})$ converges weakly in
  $\Lp{2}$.  To identify the weak limit as $u_{0}$, choose a test function
  $\varphi(x, t) = \eta(t)\,\psi(x)$ in~\eqref{eq:ns-weak} with $\psi \in
  \Ccinf(\R^{3}; \R^{3})$ divergence-free and $\eta \in \Ccinf([0, \eps))$ with
  $\eta(0) = 1$.  Letting $\eps \to 0$, the weak formulation reduces to
  $\ip{u(0^{+})}{\psi}_{\Lp{2}} = \ip{u_{0}}{\psi}_{\Lp{2}}$, confirming that
  the weak limit is $u_{0}$.  Since the limit is the same along every subsequence, the
  full net converges: $u(t) \rightharpoonup u_{0}$ as $t \to 0^{+}$.

  For the extension to $\R^{3}$, let $\Omega_{k} = B_{k}(0)$ for $k \in \N$.  For each
  $k$, the construction above yields a Leray--Hopf weak solution $u^{(k)}$ on
  $\Omega_{k} \times (0, T)$ (extending $u_{0}$ by zero outside $\Omega_{k}$ if
  $\supp u_{0}$ is compact, or projecting onto $\Lsig(\Omega_{k})$ in general).  The
  energy bounds are uniform in $k$:
  $\norm{u^{(k)}}_{\Lp{\infty}(0,T;\Lp{2})} + \norm{u^{(k)}}_{\Lp{2}(0,T;\Sobz{1})}
  \le C(\norm{u_{0}}_{\Lp{2}}, \nu)$.  A diagonal argument (extracting successive
  subsequences on $\Omega_{1}, \Omega_{2}, \ldots$ and using the Cantor diagonal
  construction) produces a limit $u$ defined on $\R^{3} \times (0, T)$ that inherits all
  four properties of \Cref{def:leray-hopf}.  The divergence-free condition passes to the
  limit because $\divergence u^{(k)} = 0$ in $\mathcal{D}'(\Omega_{k})$ for each $k$,
  and any test function $\varphi \in \Ccinf(\R^{3})$ is supported in $\Omega_{k}$ for
  $k$ sufficiently large.
\end{proof}

\subsection{Non-uniqueness and the role of the energy inequality}

\begin{remark}[Uniqueness remains open]\label{rem:uniqueness-open}
  The Leray--Hopf existence theorem does not assert uniqueness.  Whether Leray--Hopf weak
  solutions are unique for given initial data $u_{0} \in \Lsig(\R^{3})$ is one of the
  major open problems in mathematical fluid dynamics.  In two dimensions, uniqueness holds
  (the solution is in fact smooth for all time); in three dimensions, uniqueness is
  unknown.

  The energy inequality plays a selection role: it excludes ``parasitic'' weak solutions
  that spontaneously gain energy.  Without the energy inequality, non-uniqueness of weak
  solutions is known to occur.  The convex integration techniques of De Lellis and
  Sz\'{e}kelyhidi produce $\Lp{\infty}$-weak solutions of the Euler equations
  ($\nu = 0$) with prescribed energy profiles, including solutions that violate the
  energy inequality.  For $\nu > 0$, Buckmaster and Vicol (2019) constructed non-unique
  weak solutions in a regularity class strictly below the Leray--Hopf class.  The energy
  inequality is therefore not merely a technical artefact but a physically meaningful
  selection criterion.
\end{remark}

\begin{remark}[Serrin-type regularity criteria]\label{rem:serrin}
  Although uniqueness in the full Leray--Hopf class is open, additional integrability
  conditions on $u$ do imply both regularity and uniqueness.  The classical
  \emph{Serrin condition} states that if a Leray--Hopf weak solution satisfies
  $u \in \Lp{q}(0, T; \Lp{r}(\R^{3}))$ with $\frac{2}{q} + \frac{3}{r} = 1$ and
  $r > 3$, then $u$ is smooth on $(0, T]$ and is the unique Leray--Hopf solution with
  that initial data.  The Serrin condition is a \emph{scaling-critical} condition: the
  Navier--Stokes equations are invariant under the rescaling
  $u_{\lambda}(x, t) = \lambda\,u(\lambda x, \lambda^{2} t)$, and the Serrin norm
  $\norm{u}_{\Lp{q}_{t}\Lp{r}_{x}}$ is invariant under this rescaling when
  $\frac{2}{q} + \frac{3}{r} = 1$.
\end{remark}

% ---------------------------------------------------------------------------
\section{The Clay Millennium Problem}\label{sec:clay}
% ---------------------------------------------------------------------------

The existence of Leray--Hopf weak solutions guarantees that the Navier--Stokes equations
have \emph{some} kind of solution for all time.  The Clay Millennium Problem asks whether
these solutions (or, more precisely, strong solutions with smooth initial data) remain
smooth.

\subsection{The official problem statement}

The Clay Mathematics Institute formulated the Navier--Stokes existence and smoothness
problem in 2000 as one of the seven Millennium Prize Problems, with a prize of one
million dollars.  The precise mathematical statement, due to
Fefferman~\cite{fefferman2006}, poses four related problems.  We first set up the
common hypotheses and then state each problem.

Fefferman considers the Navier--Stokes equations on $\R^{3}$ with viscosity $\nu > 0$,
external force $f$, and initial velocity $u_{0}$.  A solution $(p, u)$ is called
\emph{physically reasonable} if it is smooth and has bounded energy:
\begin{equation}\label{eq:clay-smooth}
  p, u \in C^{\infty}(\R^{3} \times [0, \infty)),
\end{equation}
\begin{equation}\label{eq:clay-finite-energy}
  \int_{\R^{3}} \abs{u(x, t)}^{2} \,\dd x < C
  \qquad \text{for all } t \ge 0
  \quad \text{and some constant } C > 0.
\end{equation}
For the whole-space problems, Fefferman requires the initial data and force to be smooth
with rapid decay: for every multi-index $\alpha$, every $m \ge 0$, and every $K > 0$,
\begin{equation}\label{eq:clay-decay}
  \abs{\partial^{\alpha}_{x} u_{0}(x)}
  \;\le\;
  C_{\alpha K}\,(1 + \abs{x})^{-K}
  \qquad \text{on } \R^{3},
\end{equation}
\begin{equation}\label{eq:clay-force-decay}
  \abs{\partial^{\alpha}_{x} \partial^{m}_{t} f(x,t)}
  \;\le\;
  C_{\alpha m K}\,(1 + \abs{x} + t)^{-K}
  \qquad \text{on } \R^{3} \times [0, \infty).
\end{equation}
For the periodic problems on the torus $\R^{3}/\Z^{3}$, these decay conditions are
replaced by periodicity:
\begin{equation}\label{eq:clay-periodic}
  u_{0}(x + e_{j}) = u_{0}(x), \quad
  f(x + e_{j}, t) = f(x, t), \quad
  p(x + e_{j}, t) = p(x, t),
\end{equation}
for $1 \le j \le 3$, where $e_{j}$ are the standard unit vectors, and the solution
$u(x,t) = u(x + e_{j}, t)$ is required to be periodic as well.

With these conventions, Fefferman poses the following four problems.  A positive answer
to the Millennium Problem consists of proving either (A) or (B); a negative answer
consists of proving either (C) or (D).

\begin{conjecture}[Existence and smoothness on $\R^{3}$ (Problem A)]%
  \label{conj:clay-A}
  Take $\nu > 0$.  Let $u_{0}$ be any smooth, divergence-free vector field
  satisfying~\eqref{eq:clay-decay}, and take $f \equiv 0$.  Then there exist smooth
  functions $p(x,t)$ and $u(x,t)$ on $\R^{3} \times [0, \infty)$ satisfying the
  Navier--Stokes equations~\eqref{eq:ns-momentum}--\eqref{eq:ns-initial} together
  with~\eqref{eq:clay-smooth} and~\eqref{eq:clay-finite-energy}.
\end{conjecture}

\begin{conjecture}[Existence and smoothness on $\R^{3}/\Z^{3}$ (Problem B)]%
  \label{conj:clay-B}
  Take $\nu > 0$.  Let $u_{0}$ be any smooth, divergence-free vector field
  satisfying~\eqref{eq:clay-periodic}, and take $f \equiv 0$.  Then there exist smooth
  functions $p(x,t)$ and $u(x,t)$ on $\R^{3} \times [0, \infty)$ satisfying the
  Navier--Stokes equations~\eqref{eq:ns-momentum}--\eqref{eq:ns-initial} together
  with~\eqref{eq:clay-periodic} (for the solution),~\eqref{eq:clay-smooth},
  and~\eqref{eq:clay-finite-energy}.
\end{conjecture}

Fefferman also poses the complementary breakdown problems.

\begin{conjecture}[Breakdown on $\R^{3}$ (Problem C)]%
  \label{conj:clay-C}
  Take $\nu > 0$.  Then there exist a smooth, divergence-free vector field $u_{0}$ on
  $\R^{3}$ and a smooth force $f(x,t)$ on $\R^{3} \times [0, \infty)$,
  satisfying~\eqref{eq:clay-decay} and~\eqref{eq:clay-force-decay}, for which there exist
  no physically reasonable
  solutions~\eqref{eq:clay-smooth}--\eqref{eq:clay-finite-energy} of the Navier--Stokes
  equations on $\R^{3} \times [0, \infty)$.
\end{conjecture}

\begin{conjecture}[Breakdown on $\R^{3}/\Z^{3}$ (Problem D)]%
  \label{conj:clay-D}
  Take $\nu > 0$.  Then there exist a smooth, divergence-free vector field $u_{0}$ on
  $\R^{3}$ and a smooth force $f(x,t)$ on $\R^{3} \times [0, \infty)$,
  satisfying~\eqref{eq:clay-periodic}, for which there exist no physically reasonable
  solutions of the Navier--Stokes equations on $\R^{3} \times [0, \infty)$.
\end{conjecture}

\begin{remark}[Interpretation of the problem]
  Problems~(A) and~(B) ask two things simultaneously: \emph{existence} (there is a smooth
  solution for all time) and \emph{smoothness} (the solution does not develop
  singularities).  The decay condition~\eqref{eq:clay-decay} ensures that $u_{0}$ lies in
  every Sobolev space $\Sob{k}(\R^{3})$ and in $\Lp{p}(\R^{3})$ for all $p$, so the
  initial data is as regular as one could wish.  The bounded energy
  condition~\eqref{eq:clay-finite-energy} (note: bounded by a \emph{constant}, not merely
  finite at each time) is a natural physical requirement and connects the problem to the
  Leray--Hopf framework.

  Problems~(A) and~(B) set $f = 0$ to give solvers ``reasonable leeway''
  (Fefferman's words).  The breakdown problems~(C) and~(D) allow a nonzero force, since
  exhibiting blowup with no external forcing would be strictly harder.  A negative answer
  to the Millennium Problem via~(C) or~(D) would demonstrate initial data and forces for
  which smooth, bounded-energy solutions simply do not exist globally.
\end{remark}

\subsection{Relation to the Leray--Hopf framework}

\begin{remark}[From Leray--Hopf to the Clay Problem]\label{rem:leray-hopf-to-clay}
  The Leray--Hopf theory provides a global weak solution $u$ for any $u_{0} \in
  \Lsig(\R^{3})$, but this solution may not be smooth.  Problem~(A) asks whether,
  for smooth and rapidly decaying $u_{0}$, the weak solution is in fact a classical
  (smooth) solution.  Since smooth solutions are unique in the Leray--Hopf class (by the
  weak-strong uniqueness principle), a positive answer would imply that the Leray--Hopf
  weak solution \emph{is} the unique smooth solution.

  The connection is made precise by the following chain of implications.  If $u_{0}$ is
  smooth and rapidly decaying, classical local existence theory (Fujita--Kato, 1964)
  produces a unique smooth solution $u$ on some time interval $[0, T^{*})$.  This smooth
  local solution is also a Leray--Hopf weak solution (it satisfies all four conditions of
  \Cref{def:leray-hopf}).  Problem~(A) reduces to showing that $T^{*} = \infty$,
  i.e., the smooth local solution never breaks down.
\end{remark}

\subsection{What is known}

\begin{remark}[Partial results]\label{rem:partial-results}
  The following results are known, falling short of resolving Problem~(A):
  \begin{enumerate}[label=(\roman*)]
    \item \emph{Local existence.} For $u_{0} \in \Sob{1}(\R^{3})$ (or more generally in
      various critical spaces), there exists $T^{*} > 0$ and a unique smooth solution on
      $[0, T^{*})$.  If $\norm{u_{0}}_{\Sob{1}}$ is sufficiently small (relative to
      $\nu$), then $T^{*} = \infty$.
    \item \emph{Leray--Hopf global existence.} For $u_{0} \in \Lsig(\R^{3})$, there
      exists a global weak solution satisfying the energy inequality
      (\Cref{thm:leray-hopf-existence}).
    \item \emph{Partial regularity.} The Caffarelli--Kohn--Nirenberg theorem (1982) states
      that for any suitable Leray--Hopf weak solution, the set of singular points
      (space-time points where $u$ is not locally bounded) has one-dimensional parabolic
      Hausdorff measure zero.  In particular, the singular set cannot contain a
      space-time curve.
    \item \emph{Serrin-type criteria.} If a Leray--Hopf weak solution satisfies the
      additional integrability condition $u \in \Lp{q}(0, T; \Lp{r}(\R^{3}))$ with
      $2/q + 3/r \le 1$ and $r \ge 3$, then $u$ is smooth on $(0, T]$ (see
      \Cref{rem:serrin}).
    \item \emph{Two dimensions.} In two spatial dimensions, the Leray--Hopf weak
      solution is unique, smooth for all time, and satisfies the energy equality.  The
      three-dimensional case remains open.
  \end{enumerate}
\end{remark}

\subsection{What remains open}

\begin{remark}[The gap]\label{rem:the-gap}
  The fundamental difficulty is the \emph{supercritical} nature of the three-dimensional
  Navier--Stokes equations.  The natural energy space
  $\Lp{\infty}(0, T; \Lp{2}) \cap \Lp{2}(0, T; \Sobz{1})$ lies below the
  scaling-critical threshold: the Navier--Stokes scaling
  $u_{\lambda}(x, t) = \lambda\,u(\lambda x, \lambda^{2} t)$ preserves the
  $\Lp{3}(\R^{3})$ norm, but the energy norm $\norm{u}_{\Lp{2}}$ is not invariant
  (it scales as $\lambda^{-1/2}$).  The energy bounds are therefore too weak to control
  the nonlinearity and prevent concentration of energy at small scales.

  All known proofs of regularity (in two dimensions, or in three dimensions under
  additional hypotheses) exploit either the conservation of enstrophy (two dimensions) or
  integrability at or above the critical scaling (Serrin-type conditions).  In three
  dimensions without additional assumptions, neither mechanism is available.

  Resolving the Clay Problem requires proving one of Problems~(A)--(D) above.
  A positive answer (Problem~A or~B) would show that smooth data always produce globally
  smooth solutions.  A negative answer (Problem~C or~D) would exhibit specific data for
  which no physically reasonable solution exists globally.  As of the time of writing, neither direction has
  produced a definitive result, and the problem remains one of the deepest open questions
  in mathematical analysis.
\end{remark}

\bigskip

\noindent\textbf{Summary.}
This chapter has established the existence of Leray--Hopf weak solutions for the
three-dimensional incompressible Navier--Stokes equations with arbitrary $\Lp{2}$ initial
data (\Cref{thm:leray-hopf-existence}).  The construction proceeds via the Galerkin
method: finite-dimensional approximations using eigenfunctions of the Stokes operator
satisfy a priori energy estimates (\Cref{thm:galerkin-energy}), compactness arguments
(Aubin--Lions, Rellich--Kondrachov) extract convergent subsequences
(\Cref{thm:galerkin-convergence}), and the strong $\Lp{2}$ convergence suffices to pass
the nonlinear term to the limit (\Cref{thm:passage-limit}).  The resulting weak solution
satisfies the energy inequality (\Cref{thm:energy-inequality}) but not necessarily energy
equality.  The Leray--Hopf framework connects directly to the Clay Millennium Problem
(\Cref{conj:clay-A}): a positive resolution would show that, for smooth initial data, the
Leray--Hopf weak solution is in fact the unique smooth solution for all time.
