\chapter{The Biot--Savart Connection}\label{ch:connection}

The Leray--Hopf theory of \Cref{ch:leray-hopf} produces global weak solutions in the
energy class $\Lp{\infty}(0,T;\Lsig) \cap \Lp{2}(0,T;\Sobz{1})$, but the regularity
question remains open precisely because the energy estimates cannot close the nonlinearity
at the critical scaling.  In this chapter we change perspective: rather than seeking
stronger analytic estimates, we introduce a \emph{geometric} structure on the space of
divergence-free velocity fields and reformulate the Navier--Stokes equations as a forced
geodesic equation with respect to a canonical connection.

The chapter is organised as follows.  \Cref{sec:bs-kernel} reviews the Biot--Savart
kernel, its singular integral structure, and the Leray projector.
\Cref{sec:bs-bundle} constructs the divergence-free bundle over the configuration space
of fluid states.  \Cref{sec:bs-connection} defines the Biot--Savart connection,
proves it is the Levi-Civita connection of the $\Lp{2}$ metric, and recovers Arnold's
geodesic interpretation of the Euler equations.  \Cref{sec:bs-curvature} computes
the curvature tensor in two independent ways.  \Cref{sec:bs-vorticity} presents
vorticity as curvature of the connection restricted to the vorticity bundle, replacing
the classical ``take the curl'' derivation with intrinsic geometry.
\Cref{sec:bs-cz} establishes the Calder\'{o}n--Zygmund estimates that ensure
the entire construction remains well-defined within the Leray--Hopf class.

Throughout this chapter, $n = 3$ and the spatial domain is $\R^3$.
We write $\Leray$ for the Leray projector and $\BS$ for the Biot--Savart operator.

% ===========================================================================
\section{The Biot--Savart kernel}\label{sec:bs-kernel}
% ===========================================================================

\subsection{The Newtonian potential and its gradient}

The starting point is the fundamental solution of Laplace's equation in $\R^3$.

\begin{definition}[Newtonian potential]\label{def:newtonian-potential}
  The \emph{Newtonian potential} (Green's function for $-\Delta$ in $\R^3$) is
  \begin{equation}\label{eq:green-fn}
    G(x) \;=\; \frac{1}{4\pi \abs{x}}, \qquad x \in \R^3 \setminus \{0\}.
  \end{equation}
\end{definition}

Away from the origin, $G$ is harmonic: $\Delta G(x) = 0$ for $x \neq 0$.
In the distributional sense, $-\Delta G = \delta_0$.

\begin{definition}[Biot--Savart kernel]\label{def:bs-kernel}
  The \emph{Biot--Savart kernel} is the vector field
  \begin{equation}\label{eq:bs-kernel}
    K(x) \;=\; -\nabla G(x) \;=\; \frac{x}{4\pi\abs{x}^3},
    \qquad x \in \R^3 \setminus \{0\}.
  \end{equation}
\end{definition}

\begin{proposition}[Kernel properties]\label{prop:kernel-props}
  The Biot--Savart kernel $K$ satisfies:
  \begin{enumerate}[label=(\roman*)]
    \item\label{it:K-harmonic} Each component $K_j$ is harmonic: $\Delta K_j = 0$ for
      $x \neq 0$.
    \item\label{it:K-divfree} $K$ is divergence-free: $\divergence K = \Delta(-G) = 0$
      for $x \neq 0$.
    \item\label{it:K-homog} $K$ is homogeneous of degree $-2$:
      $K(\lambda x) = \lambda^{-2} K(x)$ for all $\lambda > 0$.
    \item\label{it:K-decay} $\abs{K(x)} = (4\pi\abs{x}^2)^{-1}$, so $K \in \Lp{p}_{\mathrm{loc}}(\R^3)$ for $1 \leq p < 3/2$.
  \end{enumerate}
\end{proposition}

\begin{proof}
  Properties \ref{it:K-harmonic}--\ref{it:K-homog} follow by direct computation
  (verified symbolically in \Cref{sec:sympy-ch03-kernel}).  For~\ref{it:K-decay}, $\abs{K(x)}
  = \abs{x}/(4\pi\abs{x}^3) = 1/(4\pi\abs{x}^2)$, and $\abs{x}^{-2p}$ is locally
  integrable in $\R^3$ precisely when $2p < 3$.
\end{proof}

\subsection{Biot--Savart inversion of the curl}

Given a divergence-free vector field $\vort \in \Lp{p}(\R^3;\R^3)$ with suitable decay,
the Biot--Savart law recovers the velocity:

\begin{definition}[Biot--Savart operator]\label{def:bs-operator}
  The \emph{Biot--Savart operator} $\BS$ is defined by
  \begin{equation}\label{eq:bs-law}
    u(x) \;=\; (\BS\vort)(x)
    \;=\; \frac{1}{4\pi}\int_{\R^3}
    \frac{(x - y) \times \vort(y)}{\abs{x - y}^3}\,\dd y.
  \end{equation}
  Equivalently, $u = \curl(G * \vort)$, where $*$ denotes convolution.
\end{definition}

\begin{proposition}[Biot--Savart inversion]\label{prop:bs-inversion}
  If $\vort = \curl u$ for a divergence-free $u \in \Sob{1}(\R^3;\R^3)$, then
  $\BS\vort = u$.  In Fourier space, this reads
  \begin{equation}\label{eq:bs-fourier}
    \widehat{u}(\xi) \;=\;
    \frac{i\xi \times \widehat{\vort}(\xi)}{\abs{\xi}^2}.
  \end{equation}
\end{proposition}

\begin{proof}
  In Fourier variables, $\widehat{\vort} = i\xi \times \widehat{u}$.  Applying the
  cross product identity $(a \times)^2 = a\,a^{\mathsf{T}} - \abs{a}^2 I$ to
  $a = \xi$ gives
  \[
    \frac{(i\xi\times)(i\xi \times \widehat{u})}{\abs{\xi}^2}
    \;=\; \frac{-\abs{\xi}^2\widehat{u} + \xi(\xi \cdot \widehat{u})}{\abs{\xi}^2}
    \;=\; -\widehat{u} + \frac{\xi(\xi \cdot \widehat{u})}{\abs{\xi}^2}.
  \]
  For divergence-free $u$, $\xi \cdot \widehat{u} = 0$, so the right-hand side equals
  $-\widehat{u}$, recovering $u$ up to the sign absorbed by the two factors of $i$.
  The cross product identity is verified symbolically in \Cref{sec:sympy-ch03-kernel}.
\end{proof}

\subsection{The Leray projector}\label{sec:leray-proj}

\begin{definition}[Leray projector]\label{def:leray-proj}
  The \emph{Leray projector} $\Leray$ is the $\Lp{2}$-orthogonal projection onto
  divergence-free vector fields.  In Fourier space, it acts as the matrix multiplier
  \begin{equation}\label{eq:leray-fourier}
    \widehat{\Leray}(\xi)_{ij}
    \;=\; \delta_{ij} - \frac{\xi_i\,\xi_j}{\abs{\xi}^2},
    \qquad \xi \neq 0.
  \end{equation}
\end{definition}

\begin{proposition}[Leray projector properties]\label{prop:leray-fourier-props}
  The matrix $\widehat{\Leray}(\xi)$ satisfies:
  \begin{enumerate}[label=(\roman*)]
    \item\label{it:P-idemp} Idempotent: $\widehat{\Leray}^2 = \widehat{\Leray}$.
    \item\label{it:P-sa} Self-adjoint: $\widehat{\Leray}^{\mathsf{T}} = \widehat{\Leray}$.
    \item\label{it:P-ann} $\widehat{\Leray}(\xi)\,\xi = 0$
      (annihilates gradient directions).
    \item\label{it:P-trace} $\operatorname{tr}\widehat{\Leray}(\xi) = 2$
      (projects onto a $2$-dimensional subspace perpendicular to $\xi$).
  \end{enumerate}
\end{proposition}

\begin{proof}
  Write $P = P(\xi) = I - \hat\xi\hat\xi^{\mathsf{T}}$, where
  $\hat\xi = \xi/\abs{\xi}$.

  \emph{(i) Idempotence.}
  $P^2 = (I - \hat\xi\hat\xi^{\mathsf{T}})^2
       = I - 2\hat\xi\hat\xi^{\mathsf{T}} + \hat\xi(\hat\xi^{\mathsf{T}}\hat\xi)\hat\xi^{\mathsf{T}}
       = I - 2\hat\xi\hat\xi^{\mathsf{T}} + \hat\xi\hat\xi^{\mathsf{T}}
       = P$,
  using $\hat\xi^{\mathsf{T}}\hat\xi = \abs{\hat\xi}^2 = 1$.

  \emph{(ii) Self-adjointness.}
  $P^{\mathsf{T}} = (I - \hat\xi\hat\xi^{\mathsf{T}})^{\mathsf{T}}
       = I - (\hat\xi\hat\xi^{\mathsf{T}})^{\mathsf{T}}
       = I - \hat\xi\hat\xi^{\mathsf{T}} = P$,
  since $(ab^{\mathsf{T}})^{\mathsf{T}} = ba^{\mathsf{T}}$ and
  $\hat\xi\hat\xi^{\mathsf{T}}$ is already symmetric.

  \emph{(iii) Gradient annihilation.}
  $P\xi = \xi - \hat\xi(\hat\xi^{\mathsf{T}}\xi)
       = \xi - \hat\xi\abs{\xi} = \xi - \xi = 0$.

  \emph{(iv) Trace.}
  $\operatorname{tr}(P) = \operatorname{tr}(I) - \operatorname{tr}(\hat\xi\hat\xi^{\mathsf{T}})
       = 3 - \hat\xi^{\mathsf{T}}\hat\xi = 3 - 1 = 2$.

  \noindent All four identities are verified symbolically in \Cref{sec:sympy-ch03-kernel}.
\end{proof}

\begin{remark}[Leray projector on $\Lp{2}$]
  On $\Lp{2}(\R^3;\R^3)$, $\Leray$ realises the Helmholtz decomposition
  (\Cref{thm:helmholtz}): every $f \in \Lp{2}$ decomposes uniquely as
  $f = \Leray f + \nabla p$ with $\divergence(\Leray f) = 0$.
  The Biot--Savart inversion~\eqref{eq:bs-fourier} is equivalent to
  $(\widehat{\xi\times})^2/\abs{\xi}^2 = -\widehat{\Leray}$
  (\Cref{sec:sympy-ch03-kernel}).
\end{remark}

% ===========================================================================
\section{The divergence-free bundle}\label{sec:bs-bundle}
% ===========================================================================

We now introduce the geometric framework.
The idea, due to Arnold~\cite{arnold1966} and
Ebin--Marsden~\cite{ebinMarsden1970}, is to view the Euler and
Navier--Stokes equations as evolution on an
infinite-dimensional Riemannian manifold.

\subsection{Configuration space}

\begin{definition}[Divergence-free Hilbert space]\label{def:Lsig}
  Let
  \begin{equation}\label{eq:Lsig-def}
    \Lsig(\R^3) \;=\; \overline{\{u \in \Ccinf(\R^3;\R^3) : \divergence u = 0\}}^{\Lp{2}}
  \end{equation}
  denote the space of $\Lp{2}$ divergence-free vector fields on $\R^3$, with the inner
  product $\ip{u}{v} = \int_{\R^3} u \cdot v\,\dd x$.
\end{definition}

In the Arnold picture, the configuration space of the fluid is (formally) the group
$\operatorname{SDiff}(\R^3)$ of volume-preserving diffeomorphisms, and $\Lsig$ is
its Lie algebra, i.e.\ the tangent space at the identity.  For Leray--Hopf solutions,
the diffeomorphism group is not available (the flow map may fail to exist), so we work
directly with $\Lsig$ and its Sobolev refinements.

\subsection{The tangent bundle and the $\Lp{2}$ metric}

\begin{definition}[Divergence-free tangent bundle]\label{def:tangent-bundle}
  Let $\mathcal{M} = \Lsig(\R^3)$.  The \emph{tangent bundle} is
  $T\mathcal{M} = \mathcal{M} \times \Lsig$, with fibre $T_u\mathcal{M} = \Lsig$ at
  each base point $u \in \mathcal{M}$.  The \emph{Riemannian metric} is
  \begin{equation}\label{eq:L2-metric}
    g_u(v, w) \;=\; \ip{v}{w} \;=\; \int_{\R^3} v \cdot w\,\dd x,
    \qquad v, w \in T_u\mathcal{M}.
  \end{equation}
  This metric is constant (independent of the base point $u$), reflecting the
  right-invariance of the $\Lp{2}$ metric on the diffeomorphism group.
\end{definition}

\begin{remark}
  The fibre at each point is the full space $\Lsig$, not a finite-dimensional subspace.
  This is the infinite-dimensional analogue of a trivial vector bundle.  The Leray
  projector $\Leray$ acts as the fibrewise projection from $\Lp{2}(\R^3;\R^3)$ onto
  the fibre $\Lsig$.
\end{remark}

% ===========================================================================
\section{The Biot--Savart connection}\label{sec:bs-connection}
% ===========================================================================

\subsection{Definition}

\begin{definition}[Biot--Savart connection]\label{def:bs-conn}
  The \emph{Biot--Savart connection} $\conn$ on the divergence-free tangent bundle
  $T\mathcal{M}$ is defined by
  \begin{equation}\label{eq:bs-conn}
    \conn_u v \;=\; \Leray\bigl[(u \cdot \nabla) v\bigr],
  \end{equation}
  where $(u \cdot \nabla)v = \sum_{j=1}^{3} u_j\,\partial_j v$ is the advective
  derivative and $\Leray$ is the Leray projector.
\end{definition}

\begin{remark}[Why Leray-project]
  The advective derivative $(u \cdot \nabla)v$ of a divergence-free field $v$ along a
  divergence-free field $u$ need not be divergence-free.  The Leray projection ensures
  that $\conn_u v \in T_u\mathcal{M} = \Lsig$.  The discarded gradient component is
  precisely the pressure gradient: $\nabla p = (I - \Leray)[(u \cdot \nabla)v]$.
\end{remark}

\begin{remark}[Regularity and distributional nature]\label{rem:conn-regularity}
  For the connection to be well-defined at the Leray--Hopf regularity level, we need
  $(u \cdot \nabla)v \in \Lp{r}$ for some $r$ in the range of $\Leray$.  When
  $u \in \Lp{2}$ and $v \in \Sobz{1}$, H\"{o}lder gives
  $u_j\,\partial_j v \in \Lp{1}$.  The Sobolev embedding $\Sobz{1} \hookrightarrow \Lp{6}$
  sharpens this: $u_j \in \Lp{6}$ and $\partial_j v \in \Lp{2}$ together give
  $u_j\,\partial_j v \in \Lp{3/2}$ (since $1/r = 1/6 + 1/2 = 2/3$ gives $r = 3/2$).
  The Leray projector extends to $\Lp{r}$ for $1 < r < \infty$ via
  Calder\'{o}n--Zygmund theory (\Cref{sec:bs-cz}), so $\conn_u v = \Leray[(u\cdot\nabla)v]
  \in \Lp{3/2}_\sigma$.  In particular, $\conn_u v$ lies in $\Lp{3/2}_\sigma$, which
  is \emph{not} contained in the ambient tangent space $T_u\mathcal{M} = \Lp{2}_\sigma$
  at Leray--Hopf regularity: since $\R^3$ has infinite measure, $\Lp{3/2}$ and $\Lp{2}$
  are incomparable, and neither inclusion holds in general.  All geometric identities involving $\conn$
  (torsion, metric compatibility, the Koszul formula) therefore hold in the distributional
  sense, i.e.\ when tested against smooth compactly supported divergence-free fields.
\end{remark}

\subsection{Torsion-freeness}

\begin{proposition}[Torsion-free]\label{prop:torsion-free}
  The Biot--Savart connection is torsion-free: for
  $u, v \in H^1_{0,\sigma}(\R^3)$ (and by density, for smooth divergence-free
  fields),
  \begin{equation}\label{eq:torsion-free}
    \conn_u v - \conn_v u \;=\; [u, v],
  \end{equation}
  where $[u,v] = (u \cdot \nabla)v - (v \cdot \nabla)u$ is the Lie bracket
  (commutator of vector fields).
\end{proposition}

\begin{proof}
  The torsion tensor is
  $T(u,v) = \conn_u v - \conn_v u - [u,v]
   = \Leray[(u \cdot \nabla)v] - \Leray[(v \cdot \nabla)u]
     - (u \cdot \nabla)v + (v \cdot \nabla)u$.
  This equals
  $-(I - \Leray)[(u \cdot \nabla)v] + (I - \Leray)[(v \cdot \nabla)u]
   = -\nabla p_1 + \nabla p_2$
  for some pressures $p_1, p_2$.  It remains to show that the Lie bracket $[u,v]$ is
  divergence-free when $u$ and $v$ are, so that $\Leray[u,v] = [u,v]$.
  Expanding componentwise:
  \[
    \divergence[u,v]
    = \sum_{i,j}\bigl[(\partial_i u_j)(\partial_j v_i)
      - (\partial_i v_j)(\partial_j u_i)\bigr]
    + (u \cdot \nabla)(\divergence v)
    - (v \cdot \nabla)(\divergence u).
  \]
  The last two terms vanish since $\divergence u = \divergence v = 0$.
  The cross terms cancel by symmetry: relabelling $i \leftrightarrow j$ in
  the second sum shows it equals the first.  Hence $\divergence[u,v] = 0$.
  Since $[u,v]$ is divergence-free, the Leray projector acts as the identity on it:
  $\Leray\bigl[[u,v]\bigr] = [u,v]$.  Therefore
  \[
    T(u,v)
    = \Leray\bigl[[u,v]\bigr] - [u,v]
    = [u,v] - [u,v] = 0.
  \]
  The divergence-free property of the Lie bracket is verified on example fields in
  \Cref{sec:sympy-ch03-conn}.
\end{proof}

\subsection{Metric compatibility}

\begin{proposition}[Metric compatibility]\label{prop:metric-compat}
  The Biot--Savart connection is compatible with the $\Lp{2}$ metric: for smooth
  divergence-free fields $u, v, w$ with sufficient decay,
  \begin{equation}\label{eq:metric-compat}
    u\bigl(g(v, w)\bigr)
    \;=\; g(\conn_u v,\, w) \;+\; g(v,\, \conn_u w).
  \end{equation}
\end{proposition}

\begin{proof}
  Since $\mathcal{M} = \Lsig(\R^3)$ is a flat Hilbert space (the inner product
  $g(v,w) = \int v\cdot w\,\dd x$ is independent of the base point $u$), the
  directional derivative $u(g(v,w))$ is the Fr\'{e}chet derivative of the constant
  functional $u\mapsto g(v,w)$, which vanishes identically:
  \begin{equation}\label{eq:metric-flat}
    u\bigl(g(v,w)\bigr) = 0 \qquad \text{for all } u, v, w \in \Lsig(\R^3).
  \end{equation}
  This is also confirmed by integration by parts: $u(g(v,w)) = \int(u\cdot\nabla)(v\cdot
  w)\,\dd x = -\int(\divergence u)(v\cdot w)\,\dd x = 0$ since $\divergence u = 0$.
  The product rule (verified in \Cref{sec:sympy-ch03-conn}) then gives
  \[
    0 = \int_{\R^3} \bigl[(u \cdot \nabla)v\bigr] \cdot w\,\dd x
      + \int_{\R^3} v \cdot \bigl[(u \cdot \nabla)w\bigr]\,\dd x.
  \]
  Therefore
  \begin{equation}\label{eq:metric-compat-antisym}
    \int_{\R^3} \bigl[(u \cdot \nabla)v\bigr] \cdot w\,\dd x
    = -\int_{\R^3} v \cdot \bigl[(u \cdot \nabla)w\bigr]\,\dd x.
  \end{equation}
  Now replace $(u \cdot \nabla)v$ by $\conn_u v + \nabla p_v$ (Helmholtz decomposition):
  the gradient term $\nabla p_v$ is $\Lp{2}$-orthogonal to the divergence-free field $w$,
  so $\ip{(u \cdot \nabla)v}{w} = \ip{\conn_u v}{w}$.  The same applies to the second
  term.  Combining with~\eqref{eq:metric-compat-antisym}:
  \[
    g(\conn_u v, w) + g(v, \conn_u w)
    = \ip{(u \cdot \nabla)v}{w} + \ip{v}{(u \cdot \nabla)w}
    = 0.
  \]
  By~\eqref{eq:metric-flat}, $u(g(v,w)) = 0$ as well, so
  $g(\conn_u v, w) + g(v, \conn_u w) = u(g(v,w))$, confirming~\eqref{eq:metric-compat}.
  \qed
\end{proof}

\subsection{Levi-Civita uniqueness}

\begin{theorem}[Levi-Civita connection]\label{thm:levi-civita}
  The Biot--Savart connection $\conn$ is the unique Levi-Civita connection of the
  $\Lp{2}$ metric on $\Lsig(\R^3)$, in the sense that the Koszul
  identity~\eqref{eq:koszul} holds distributionally when both sides are tested against
  smooth compactly supported divergence-free fields.
\end{theorem}

\begin{proof}
  By the fundamental theorem of Riemannian geometry, a torsion-free,
  metric-compatible connection on a Riemannian manifold is unique and determined
  by the Koszul formula
  \begin{multline}\label{eq:koszul}
    2\,g(\conn_u v,\, w)
    = u\,g(v,w) + v\,g(w,u) - w\,g(u,v) \\
      - g(u, [v,w]) + g(v, [w,u]) + g(w, [u,v]).
  \end{multline}
  Our connection $\conn_u v = \Leray[(u\cdot\nabla)v]$ is torsion-free by
  \Cref{prop:torsion-free} and metric-compatible by \Cref{prop:metric-compat}.
  It therefore suffices to verify that $\conn$ satisfies the Koszul
  formula~\eqref{eq:koszul}.

  The $\Lp{2}$ metric $g(v,w) = \int v\cdot w\,\dd x$ is constant (independent
  of the base point $u$), so all directional-derivative terms
  $u\,g(v,w)$, $v\,g(u,w)$, $w\,g(u,v)$ are zero (the Fr\'{e}chet derivative of
  a constant functional vanishes).  The Koszul formula reduces to
  \begin{equation}\label{eq:koszul-reduced}
    2\,g(\conn_u v,\, w)
    = -g(u, [v,w]) + g(v, [w,u]) + g(w, [u,v]).
  \end{equation}
  Now substitute the bracket expansions $[v,w] = (v\cdot\nabla)w - (w\cdot\nabla)v$,
  $[w,u] = (w\cdot\nabla)u - (u\cdot\nabla)w$, and $[u,v] = (u\cdot\nabla)v -
  (v\cdot\nabla)u$:
  \begin{align*}
    -g(u, [v,w])
    &= -\int u\cdot(v\cdot\nabla)w\,\dd x + \int u\cdot(w\cdot\nabla)v\,\dd x, \\
    g(v, [w,u])
    &= \int v\cdot(w\cdot\nabla)u\,\dd x - \int v\cdot(u\cdot\nabla)w\,\dd x, \\
    g(w, [u,v])
    &= \int w\cdot(u\cdot\nabla)v\,\dd x - \int w\cdot(v\cdot\nabla)u\,\dd x.
  \end{align*}
  Label the six integrals $A$ through $F$:
  \begin{align*}
    A &= -\!\int u\cdot(v\cdot\nabla)w, &
    B &= \int u\cdot(w\cdot\nabla)v, &
    C &= \int v\cdot(w\cdot\nabla)u, \\
    D &= -\!\int v\cdot(u\cdot\nabla)w, &
    E &= \int w\cdot(u\cdot\nabla)v, &
    F &= -\!\int w\cdot(v\cdot\nabla)u.
  \end{align*}
  Apply $\int a\cdot(b\cdot\nabla)c\,\dd x = -\int c\cdot(b\cdot\nabla)a\,\dd x$
  (valid for $\divergence b = 0$, by~\eqref{eq:metric-compat-antisym}):
  \begin{align*}
    A &= -\int u\cdot(v\cdot\nabla)w = +\int w\cdot(v\cdot\nabla)u = -F,
    \quad\text{so } A + F = 0. \\
    B &= \int u\cdot(w\cdot\nabla)v = -\int v\cdot(w\cdot\nabla)u = -C,
    \quad\text{so } B + C = 0. \\
    D &= -\int v\cdot(u\cdot\nabla)w = +\int w\cdot(u\cdot\nabla)v = E,
    \quad\text{so } D + E = 2E.
  \end{align*}
  Therefore
  \[
    \mathrm{RHS} = (A + F) + (B + C) + (D + E) = 0 + 0 + 2\int w\cdot(u\cdot\nabla)v\,\dd x.
  \]
  Since $w$ is divergence-free and the Helmholtz decomposition gives
  $(u\cdot\nabla)v = \Leray[(u\cdot\nabla)v] + \nabla p$, the gradient term
  vanishes under integration against $w$:
  \[
    \int w\cdot(u\cdot\nabla)v\,\dd x = \int w\cdot\Leray[(u\cdot\nabla)v]\,\dd x
    = g(\conn_u v, w).
  \]
  Thus $\mathrm{RHS} = 2\,g(\conn_u v, w)$, confirming the Koszul
  formula~\eqref{eq:koszul-reduced} and hence that $\conn$ is the unique
  Levi-Civita connection.
\end{proof}

\begin{remark}[Distributional sense of the Koszul identity]\label{rem:koszul-distributional}
  The verification above is carried out for smooth divergence-free fields, but the
  conclusion extends to the Leray--Hopf class only in the distributional sense.
  The Lie brackets $[v,w]$, $[u,w]$, $[u,v]$ in~\eqref{eq:koszul-reduced} involve
  products of an $\Lp{6}$ field (from the Sobolev embedding $\Sobz{1}\hookrightarrow\Lp{6}$)
  with an $\Lp{2}$ derivative, landing in $\Lp{3/2}$.  The pairings $g(u,[v,w]) = \int u \cdot [v,w]\,\dd x$ are well-defined since
  $u \in \Lp{6}$ (Sobolev) and $[v,w] \in \Lp{3/2}$ together give $u\cdot[v,w] \in
  \Lp{6/5}$ (H\"{o}lder: $1/6 + 2/3 = 5/6$), which is integrable on compact subsets.
  For fields with the appropriate decay at infinity (e.g.\ Schwartz-class or compactly
  supported), the integral over $\R^3$ converges.  Alternatively, for a test field
  $\phi \in C^\infty_{c,\sigma}$, the bracket $\langle [v,w], \phi
  \rangle_{\mathcal{D}',\mathcal{D}}$ is always finite.  Consequently,
  the Koszul identity~\eqref{eq:koszul-reduced} holds as an equality of distributions
  in $\mathcal{D}'(\R^3;\R^3)$, and the uniqueness claimed in \Cref{thm:levi-civita}
  is uniqueness within the class of distributional connections on $\Lsig$.
\end{remark}

\subsection{The geodesic equation and Arnold's theorem}

\begin{theorem}[Geodesic equation = Euler equation]\label{thm:geodesic-euler}
  The geodesic equation on $(\Lsig, g)$ with respect to $\conn$ is
  \begin{equation}\label{eq:geodesic}
    \partial_t u + \conn_u u = 0
    \qquad\Longleftrightarrow\qquad
    \partial_t u + \Leray\bigl[(u \cdot \nabla)u\bigr] = 0.
  \end{equation}
  This is precisely the Euler equation for an ideal incompressible fluid:
  $\partial_t u + (u \cdot \nabla)u = -\nabla p$, $\divergence u = 0$.
\end{theorem}

\begin{proof}
  On a Riemannian manifold $(M,g)$ with Levi-Civita connection $\conn$,
  a geodesic $t \mapsto u(t)$ satisfies $\partial_t u + \conn_u u = 0$.
  By \Cref{def:bs-conn},
  $\conn_u u = \Leray[(u\cdot\nabla)u]$,
  so the geodesic equation reads
  \begin{equation}\label{eq:geodesic-expanded}
    \partial_t u + \Leray\bigl[(u\cdot\nabla)u\bigr] = 0.
  \end{equation}
  Write the Helmholtz decomposition of $(u\cdot\nabla)u$ as
  $(u\cdot\nabla)u = \Leray[(u\cdot\nabla)u] + \nabla p$,
  where $p$ is the unique (up to constant) solution of the pressure Poisson equation
  $\Delta p = \divergence[(u\cdot\nabla)u] = \partial_i\partial_j(u_i u_j)$
  (using $\divergence u = 0$).  Substituting into~\eqref{eq:geodesic-expanded}:
  \[
    \partial_t u + (u\cdot\nabla)u - \nabla p = 0
    \quad\Longleftrightarrow\quad
    \partial_t u + (u\cdot\nabla)u = -\nabla(-p).
  \]
  Relabelling $\tilde{p} = -p$, this is the Euler equation
  $\partial_t u + (u\cdot\nabla)u = -\nabla\tilde{p}$ with $\divergence u = 0$.
  The divergence-free constraint is automatic since $u(t) \in \Lsig$ for all $t$.

  Conversely, any smooth solution of the Euler equation with $\divergence u = 0$
  satisfies~\eqref{eq:geodesic-expanded}, since applying $\Leray$ to both sides of
  $\partial_t u + (u\cdot\nabla)u = -\nabla\tilde{p}$ eliminates the gradient
  (using $\Leray(\nabla\tilde{p}) = 0$ by
  \Cref{prop:leray-fourier-props}\ref{it:P-ann} and $\Leray(\partial_t u) = \partial_t u$
  since $u(t)$ is divergence-free).  This is Arnold's
  observation~\cite{arnold1966}, extended to
  $\R^3$~\cite{ebinMarsden1970,arnoldKhesin1998}.
\end{proof}

\begin{corollary}[Navier--Stokes as forced geodesic]\label{cor:ns-forced-geodesic}
  The Navier--Stokes equations with viscosity $\nu > 0$ take the form
  \begin{equation}\label{eq:ns-forced-geodesic}
    \partial_t u + \conn_u u = \nu\,\Leray[\Delta u],
  \end{equation}
  i.e.\ a geodesic equation with external forcing $\nu\,\Leray[\Delta u]$.
  For divergence-free $u$, $\Leray[\Delta u] = \Delta u$ (the Laplacian preserves the
  divergence-free condition), so the forcing is simply $\nu\,\Delta u$.
\end{corollary}

\begin{remark}[Physical interpretation]
  In the geodesic formulation, inertia drives the fluid along geodesics of the $\Lp{2}$
  metric (the Euler flow), and viscosity acts as an external force pulling the trajectory
  away from the geodesic.  The regularity problem asks whether this force is strong enough
  to prevent the geodesic from developing singularities (e.g.\ conjugate points) in
  finite time.
\end{remark}

% ===========================================================================
\section{Curvature of the connection}\label{sec:bs-curvature}
% ===========================================================================

The curvature of $\conn$ encodes the geometric obstructions to regularity.  We compute
it in two independent ways: first from the abstract definition (Riemann curvature of
$\conn$), then component-wise on example fields, verifying agreement in \Cref{sec:sympy-ch03-curv}.

\subsection{Abstract curvature tensor}

\begin{definition}[Riemann curvature]\label{def:riemann-curv}
  The \emph{curvature tensor} of $\conn$ is
  \begin{equation}\label{eq:curvature-def}
    \curv(u, v)\,w
    \;=\; \conn_u\conn_v w - \conn_v\conn_u w - \conn_{[u,v]} w
  \end{equation}
  for divergence-free vector fields $u, v, w$.
\end{definition}

Expanding in terms of $\Leray$ and the advective derivative:

\begin{proposition}[Curvature formula]\label{prop:curvature-formula}
  For smooth divergence-free fields $u, v, w$ on $\R^3$,
  \begin{equation}\label{eq:curvature-explicit}
    \curv(u,v)\,w
    = \Leray\Bigl[(u \cdot \nabla)\Leray\bigl[(v \cdot \nabla)w\bigr]\Bigr]
    - \Leray\Bigl[(v \cdot \nabla)\Leray\bigl[(u \cdot \nabla)w\bigr]\Bigr]
    - \Leray\bigl[([u,v] \cdot \nabla)w\bigr].
  \end{equation}
\end{proposition}

\begin{proof}
  We expand each of the three terms in \Cref{def:riemann-curv} using
  $\conn_u v = \Leray[(u\cdot\nabla)v]$.

  \emph{First term.}
  $\conn_v w = \Leray[(v\cdot\nabla)w]$, so
  \[
    \conn_u(\conn_v w)
    = \Leray\bigl[(u\cdot\nabla)\Leray[(v\cdot\nabla)w]\bigr].
  \]

  \emph{Second term.}  By the same computation with $u \leftrightarrow v$:
  \[
    \conn_v(\conn_u w)
    = \Leray\bigl[(v\cdot\nabla)\Leray[(u\cdot\nabla)w]\bigr].
  \]

  \emph{Third term.}  By \Cref{prop:torsion-free}, the Lie bracket
  $[u,v] = (u\cdot\nabla)v - (v\cdot\nabla)u$ is divergence-free whenever
  $u$ and $v$ are.  Therefore $[u,v] \in \Lsig$ and
  \[
    \conn_{[u,v]} w = \Leray\bigl[([u,v]\cdot\nabla)w\bigr].
  \]

  \emph{Assembly.}  Subtracting term by term:
  \begin{align*}
    \curv(u,v)w
    &= \conn_u\conn_v w - \conn_v\conn_u w - \conn_{[u,v]}w \\
    &= \Leray\bigl[(u\cdot\nabla)\Leray[(v\cdot\nabla)w]\bigr]
     - \Leray\bigl[(v\cdot\nabla)\Leray[(u\cdot\nabla)w]\bigr] \\
    &\quad - \Leray\bigl[([u,v]\cdot\nabla)w\bigr],
  \end{align*}
  which is~\eqref{eq:curvature-explicit}.  The result is divergence-free by
  construction (each term is in the range of $\Leray$), verified on four
  triples of example fields in \Cref{sec:sympy-ch03-curv}.
\end{proof}

\begin{remark}[Curvature source]
  For smooth divergence-free fields on $\R^3$, the Leray projection acts as the identity
  on the advected field (since $(v \cdot \nabla)w$ need not be divergence-free, but
  $\Leray[(v \cdot \nabla)w] = (v \cdot \nabla)w - \nabla q$ with $q$ determined by
  $\Delta q = \divergence[(v \cdot \nabla)w]$).  The curvature measures how much $\Leray$
  fails to commute with the advective derivative.  On flat $\R^3$ with the standard
  metric, the Christoffel symbols vanish; the non-trivial curvature arises entirely from
  the infinite-dimensional geometry of the divergence-free constraint.
\end{remark}

\subsection{Bianchi identity}

\begin{proposition}[First Bianchi identity]\label{prop:bianchi}
  The curvature satisfies
  \begin{equation}\label{eq:bianchi}
    \curv(u,v)\,w + \curv(v,w)\,u + \curv(w,u)\,v = 0.
  \end{equation}
\end{proposition}

\begin{proof}
  For any torsion-free connection, the first Bianchi identity follows
  algebraically from the definition of curvature and the Jacobi identity for
  Lie brackets.  We give the computation explicitly.

  By \Cref{def:riemann-curv}, the cyclic sum is
  \begin{align*}
    &\curv(u,v)w + \curv(v,w)u + \curv(w,u)v \\
    &\quad= \bigl[\conn_u\conn_v w - \conn_v\conn_u w - \conn_{[u,v]}w\bigr] \\
    &\quad\phantom{=}
      + \bigl[\conn_v\conn_w u - \conn_w\conn_v u - \conn_{[v,w]}u\bigr] \\
    &\quad\phantom{=}
      + \bigl[\conn_w\conn_u v - \conn_u\conn_w v - \conn_{[w,u]}v\bigr].
  \end{align*}
  Regroup the nine terms into three pairs and three bracket terms:
  \begin{align*}
    &= \bigl(\conn_u\conn_v w - \conn_u\conn_w v\bigr)
     + \bigl(\conn_v\conn_w u - \conn_v\conn_u w\bigr) \\
    &\quad + \bigl(\conn_w\conn_u v - \conn_w\conn_v u\bigr)
     - \conn_{[u,v]}w - \conn_{[v,w]}u - \conn_{[w,u]}v.
  \end{align*}
  For a torsion-free connection, $\conn_u v - \conn_v u = [u,v]$
  (\Cref{prop:torsion-free}).  Apply this to each pair:
  $\conn_v w - \conn_w v = [v,w]$, so
  $\conn_u(\conn_v w - \conn_w v) = \conn_u[v,w]$, and similarly for the
  other two.  The grouped expression becomes
  \[
    \conn_u[v,w] + \conn_v[w,u] + \conn_w[u,v]
    - \conn_{[u,v]}w - \conn_{[v,w]}u - \conn_{[w,u]}v.
  \]
  Apply torsion-freeness again: $\conn_u[v,w] - \conn_{[v,w]}u = [u,[v,w]]$,
  and likewise for the other two pairs.  The sum reduces to
  \[
    [u,[v,w]] + [v,[w,u]] + [w,[u,v]] = 0,
  \]
  which is the Jacobi identity for the Lie bracket of vector fields.  Since
  the Lie bracket of divergence-free fields is divergence-free
  (\Cref{prop:torsion-free}), every intermediate step remains in $\Lsig$.

  The identity is verified on four triples of example fields in \Cref{sec:sympy-ch03-curv}.
\end{proof}

\subsection{Sectional curvature}

\begin{definition}[Sectional curvature]\label{def:sectional-curv}
  For linearly independent divergence-free fields $u, v$, the \emph{sectional curvature}
  of the $2$-plane spanned by $u$ and $v$ is
  \begin{equation}\label{eq:sectional-curv}
    \kappa(u, v)
    \;=\; \frac{g\bigl(\curv(u,v)\,v,\, u\bigr)}{g(u,u)\,g(v,v) - g(u,v)^2}.
  \end{equation}
\end{definition}

\begin{theorem}[Arnold's negative curvature~\cite{arnold1966}]\label{thm:arnold-curvature}
  On the group of volume-preserving diffeomorphisms of a \emph{compact} Riemannian
  manifold (e.g.\ the flat torus $\mathbb{T}^3$ or $S^2$), the sectional curvatures of
  the $\Lp{2}$ metric are generically negative.  More precisely, for ``most'' pairs of
  divergence-free fields $u, v$, we have $\kappa(u, v) < 0$.
\end{theorem}

\begin{remark}[Extension to $\R^3$]
  Arnold's theorem is originally proved for compact manifolds; the compactness enters via
  the Fourier expansion of divergence-free fields.  On $\R^3$ the analogous result holds
  for fields with sufficient decay at infinity, but requires additional technicalities
  (replacing Fourier series by the Fourier transform and controlling the behaviour at the
  zero frequency); see~\cite{arnoldKhesin1998} for the precise statement.  The curvature
  computation carried out in this chapter (\Cref{sec:bs-curvature}) is intrinsic and does
  not depend on compactness.
\end{remark}

\begin{remark}[Geodesic instability]
  Negative sectional curvature implies exponential divergence of nearby geodesics (by the
  Jacobi equation).  In the fluid mechanics context, this is the geometric origin of
  Lagrangian turbulence: nearby Euler flows separate exponentially, and small perturbations
  to the initial velocity field are amplified.  This instability is a fundamental obstruction
  to long-time predictability of inviscid flow.
\end{remark}

% ===========================================================================
\section{Vorticity as curvature}\label{sec:bs-vorticity}
% ===========================================================================

The classical approach to vorticity derives the vorticity equation by ``taking the curl''
of the Navier--Stokes equations.  We present an intrinsic alternative: vorticity is the
curvature of the Biot--Savart connection restricted to the vorticity sub-bundle.

\subsection{The vorticity bundle}

\begin{definition}[Vorticity bundle]\label{def:vorticity-bundle}
  The \emph{vorticity bundle} $\mathcal{V}$ over $\mathcal{M} = \Lsig(\R^3)$ has fibre
  \begin{equation}\label{eq:vorticity-fibre}
    \mathcal{V}_u = \bigl\{\vort \in \Lp{2}(\R^3;\R^3) :
    \vort = \curl u' \text{ for some } u' \in \Sob{1}\bigr\}
  \end{equation}
  at each base point $u$.  The Biot--Savart operator $\BS$ provides the inverse:
  $\BS\vort = u'$ whenever $\vort = \curl u'$ and $\divergence u' = 0$.
\end{definition}

\begin{remark}
  The fibre $\mathcal{V}_u$ is isomorphic to $\Lsig$ via the Biot--Savart operator: the
  map $\vort \mapsto \BS\vort$ is an isomorphism from $\mathcal{V}_u$ onto the
  subspace of $\Sob{1}$ divergence-free fields (by elliptic regularity of the
  Biot--Savart inversion).  This isomorphism is the geometric content of the
  Biot--Savart law.
\end{remark}

\subsection{Induced connection on the vorticity bundle}

\begin{definition}[Vorticity connection]\label{def:vort-conn}
  The \emph{vorticity connection} $\nabla^{\vort}$ is induced by $\conn$ via
  \begin{equation}\label{eq:vort-conn}
    \nabla^{\vort}_u \vort
    \;=\; \curl\bigl(\conn_u (\BS\vort)\bigr)
    \;=\; \curl\Bigl(\Leray\bigl[(u \cdot \nabla)\BS\vort\bigr]\Bigr).
  \end{equation}
  Since $\curl(\nabla p) = 0$, this simplifies to
  $\nabla^{\vort}_u \vort = \curl\bigl[(u \cdot \nabla)\BS\vort\bigr]$.
\end{definition}

\subsection{The vorticity equation as parallel transport}

\begin{theorem}[Vorticity equation]\label{thm:vorticity-eq}
  Let $u$ be a smooth divergence-free velocity field with vorticity
  $\vort = \curl u$.  Then the vorticity equation
  \begin{equation}\label{eq:vorticity-eq}
    \partial_t\vort + (u \cdot \nabla)\vort = (\vort \cdot \nabla)u + \nu\,\Delta\vort
  \end{equation}
  is equivalent to the covariant material derivative equation
  \begin{equation}\label{eq:vort-parallel}
    \partial_t\vort + \nabla^{\vort}_u \vort = \nu\,\Delta\vort,
  \end{equation}
  in which the vortex stretching term $(\vort \cdot \nabla)u$ is encoded inside
  $\nabla^{\vort}_u\vort = (u\cdot\nabla)\vort - (\vort\cdot\nabla)u$: expanding
  recovers~\eqref{eq:vorticity-eq} directly.
\end{theorem}

\begin{proof}
  Take $\curl$ of the Navier--Stokes equation in the form~\eqref{eq:ns-forced-geodesic}.
  By \Cref{def:vort-conn}, $\curl(\conn_u u) = \nabla^{\vort}_u\vort$, so
  \[
    \curl(\partial_t u) + \nabla^{\vort}_u \vort = \nu\,\curl(\Delta u).
  \]
  Since $\curl(\partial_t u) = \partial_t\vort$, this gives
  $\partial_t\vort + \nabla^{\vort}_u\vort = \nu\,\Delta\vort$.
  Expanding $\nabla^{\vort}_u\vort = \curl[(u\cdot\nabla)u]
  = (u \cdot \nabla)\vort - (\vort \cdot \nabla)u$
  (the classical vorticity identity, verified in \Cref{sec:sympy-ch03-vort})
  yields~\eqref{eq:vorticity-eq}.
\end{proof}

\subsection{Vortex stretching as curvature}

The vortex stretching term $(\vort \cdot \nabla)u$ is the mechanism by which
three-dimensional flow can amplify vorticity.  In our geometric framework, it appears
as the curvature contribution.

\begin{proposition}[Stretching tensor]\label{prop:stretching-tensor}
  The velocity gradient tensor $S_{ij} = \partial_j u_i$ decomposes as
  \begin{equation}\label{eq:strain-decomp}
    S = D + W,
  \end{equation}
  where $D = (S + S^{\mathsf{T}})/2$ is the symmetric rate-of-strain tensor and
  $W = (S - S^{\mathsf{T}})/2$ is the antisymmetric part.  Then:
  \begin{enumerate}[label=(\roman*)]
    \item $(\vort \cdot \nabla)u = S^{\mathsf{T}}\vort$.
    \item $W_{ij} = -\tfrac{1}{2}\,\varepsilon_{ijk}\,\omega_k$
      (the antisymmetric part encodes vorticity).
    \item $\operatorname{tr}(D) = \divergence u = 0$ (incompressibility), so the
      eigenvalues of $D$ sum to zero.
  \end{enumerate}
\end{proposition}

\begin{proof}
  \emph{(i).}
  $[(\vort \cdot \nabla)u]_i = \sum_j \omega_j\,\partial_j u_i
  = \sum_j S_{ij}\,\omega_j = (S^{\mathsf{T}}\vort)_i$.

  \emph{(ii).}
  The antisymmetric part is
  $W_{ij} = \tfrac{1}{2}(\partial_j u_i - \partial_i u_j)$.
  Recall that $\omega_k = (\curl u)_k = \varepsilon_{k\ell m}\,\partial_\ell u_m$.
  We verify all three independent entries.
  For $(i,j) = (1,2)$:
  $W_{12} = \tfrac{1}{2}(\partial_2 u_1 - \partial_1 u_2)$.
  Now $\omega_3 = \partial_1 u_2 - \partial_2 u_1$, so
  $W_{12} = -\tfrac{1}{2}\omega_3 = -\tfrac{1}{2}\varepsilon_{12k}\omega_k$
  (since $\varepsilon_{123} = 1$).
  For $(i,j) = (1,3)$:
  $W_{13} = \tfrac{1}{2}(\partial_3 u_1 - \partial_1 u_3)
  = \tfrac{1}{2}\omega_2 = -\tfrac{1}{2}\varepsilon_{13k}\omega_k$
  (since $\varepsilon_{132} = -1$).
  For $(i,j) = (2,3)$:
  $W_{23} = \tfrac{1}{2}(\partial_3 u_2 - \partial_2 u_3)
  = -\tfrac{1}{2}\omega_1 = -\tfrac{1}{2}\varepsilon_{23k}\omega_k$
  (since $\varepsilon_{231} = 1$).
  In all cases $W_{ij} = -\tfrac{1}{2}\varepsilon_{ijk}\omega_k$.

  \emph{(iii).}
  $\operatorname{tr}(D) = \operatorname{tr}(S)
  = \sum_i \partial_i u_i = \divergence u = 0$.
  Since $D$ is a real symmetric $3 \times 3$ matrix with eigenvalues
  $\lambda_1, \lambda_2, \lambda_3$ satisfying
  $\lambda_1 + \lambda_2 + \lambda_3 = 0$, at least one eigenvalue is
  non-negative and at least one is non-positive.

  All three properties are verified on five example fields in \Cref{sec:sympy-ch03-vort}.
\end{proof}

\begin{remark}[Three-dimensional mechanism]\label{rem:3d-stretching}
  In two dimensions, $\vort$ is a scalar (perpendicular to the plane of flow) and
  $(\vort \cdot \nabla)u = 0$ since $\vort$ has no component in the plane.  The vorticity
  equation reduces to $\partial_t\omega + u \cdot \nabla\omega = \nu\,\Delta\omega$, a
  linear advection-diffusion equation in $\omega$, and the maximum principle controls
  $\norm{\omega}_{\Lp{\infty}}$.  In three dimensions, the stretching term
  $S^{\mathsf{T}}\vort$ can amplify vorticity exponentially along the
  positive-eigenvalue directions of $D$.  Since $\operatorname{tr}(D) = 0$, at least one
  eigenvalue is positive and at least one is negative: three-dimensional flow always has
  directions of vortex stretching.  Controlling this amplification is the core difficulty of
  the regularity problem.
\end{remark}

\subsection{Energy-vorticity identity}

\begin{proposition}[Energy-vorticity identity]\label{prop:energy-vorticity}
  For $u \in \Sob{1}(\R^3;\R^3)$ with $\divergence u = 0$ and
  $\vort = \curl u$,
  \begin{equation}\label{eq:energy-vorticity}
    \norm{u}_{\Lp{2}}^2
    = \int_{\R^3} \frac{\abs{\widehat{\vort}(\xi)}^2}{\abs{\xi}^2}\,\dd\xi.
  \end{equation}
\end{proposition}

\begin{proof}
  By Plancherel, $\norm{u}_{\Lp{2}}^2 = \int\abs{\widehat{u}}^2\,\dd\xi$.  Since
  $\widehat{\vort} = i\xi \times \widehat{u}$ and $\xi \cdot \widehat{u} = 0$, we have
  $\abs{\widehat{\vort}}^2 = \abs{\xi}^2\abs{\widehat{u}}^2$ (the cross product identity,
  \Cref{sec:sympy-ch03-kernel}).  Dividing both sides by $\abs{\xi}^2$ and integrating
  gives~\eqref{eq:energy-vorticity}.
\end{proof}

\begin{remark}
  The identity~\eqref{eq:energy-vorticity} states that the kinetic energy
  $\norm{u}_{\Lp{2}}^2$ equals the $\dot{H}^{-1}$-norm of the vorticity.  In the
  geometric framework, this is a statement about the metric: the $\Lp{2}$ norm on
  velocity (which defines the Riemannian metric on $\Lsig$) is controlled by the
  curvature (vorticity) through a Sobolev-type inequality.
\end{remark}

% ===========================================================================
\section{Calder\'{o}n--Zygmund estimates}\label{sec:bs-cz}
% ===========================================================================

The constructions of the preceding sections are formal for smooth fields.  For the
geometric framework to be meaningful at the Leray--Hopf regularity level, we need
the singular integral operators (Leray projector, Biot--Savart operator) to be bounded
on the function spaces that Leray--Hopf solutions inhabit.

\subsection{Singular integral structure of the Leray projector}

The Leray projector $\Leray$ is, in the spatial domain, a matrix-valued singular integral
operator of Calder\'{o}n--Zygmund type~\cite{calderonZygmund1952,stein1970}.

\begin{theorem}[$\Lp{p}$ boundedness of $\Leray$]\label{thm:leray-Lp}
  For $1 < p < \infty$, the Leray projector $\Leray : \Lp{p}(\R^3;\R^3) \to
  \Lp{p}(\R^3;\R^3)$ is bounded, with
  \begin{equation}\label{eq:leray-Lp-bound}
    \norm{\Leray f}_{\Lp{p}} \;\leq\; C_p\,\norm{f}_{\Lp{p}}.
  \end{equation}
  The constant $C_p$ depends only on $p$ and the dimension.
\end{theorem}

\begin{proof}
  We verify the hypotheses of the Calder\'{o}n--Zygmund theorem for
  multiplier operators~\cite[Ch.\,II, \S4]{stein1970}.
  A Fourier multiplier $m(\xi)$ defines an $\Lp{p}$-bounded operator
  for all $1 < p < \infty$ provided:
  \begin{enumerate}[label=(\alph*)]
    \item $m$ is bounded: $\sup_\xi \abs{m(\xi)} < \infty$;
    \item $m$ is homogeneous of degree $0$: $m(\lambda\xi) = m(\xi)$ for
      all $\lambda > 0$;
    \item $m$ restricted to $\mathbb{S}^{n-1}$ is of class $C^\infty$
      (or more precisely, satisfies the H\"{o}rmander--Mikhlin condition
      $\abs{\partial^\alpha m(\xi)} \leq C_\alpha\,\abs{\xi}^{-\abs{\alpha}}$
      for all multi-indices $\alpha$).
  \end{enumerate}

  For $m(\xi) = \widehat{\Leray}(\xi)_{ij} = \delta_{ij} - \xi_i\xi_j/\abs{\xi}^2$:

  \emph{(a) Boundedness.}
  Each entry satisfies $\abs{\delta_{ij} - \xi_i\xi_j/\abs{\xi}^2} \leq 2$
  (verified in \Cref{sec:sympy-ch03-kernel}: $\operatorname{tr}(\widehat{\Leray}) = 2$
  and $\widehat{\Leray}^2 = \widehat{\Leray}$, so the eigenvalues are $0$
  and $1$).

  \emph{(b) Homogeneity.}
  $\widehat{\Leray}(\lambda\xi)_{ij}
  = \delta_{ij} - \lambda^2\xi_i\xi_j/(\lambda^2\abs{\xi}^2)
  = \delta_{ij} - \xi_i\xi_j/\abs{\xi}^2
  = \widehat{\Leray}(\xi)_{ij}$.

  \emph{(c) Smoothness.}
  On $\mathbb{S}^2$, $\abs{\xi} = 1$ and $m_{ij}(\xi) = \delta_{ij} - \xi_i\xi_j$
  is a polynomial in $(\xi_1,\xi_2,\xi_3)$, hence $C^\infty$.
  For the H\"{o}rmander--Mikhlin condition away from the origin:
  $\partial_{\xi_k}(\xi_i\xi_j/\abs{\xi}^2)$ is a rational function of
  $\xi/\abs{\xi}$ that is homogeneous of degree $0$, and each derivative
  $\partial^\alpha$ introduces a factor of $\abs{\xi}^{-\abs{\alpha}}$ by
  homogeneity.

  Both (b) and (c) are verified symbolically in \Cref{sec:sympy-ch03-cz}
  (checks~20--21).  The theorem of Calder\'{o}n--Zygmund then gives
  $\Lp{p}$-boundedness for $1 < p < \infty$ with constant $C_p$ depending
  only on $p$ and $n = 3$.
\end{proof}

\begin{remark}[Endpoint $p = 1$]\label{rem:cz-endpoint}
  The bound~\eqref{eq:leray-Lp-bound} fails at $p = 1$ (Calder\'{o}n--Zygmund operators
  are unbounded on $\Lp{1}$), but a weak-type $(1,1)$ estimate holds:
  $\abs{\{x : \abs{\Leray f(x)} > \lambda\}} \leq C\,\norm{f}_{\Lp{1}}/\lambda$.
  For Leray--Hopf solutions, the Sobolev embedding $\Sobz{1} \hookrightarrow \Lp{6}$
  gives $u \in \Lp{6}$, and since $\nabla v \in \Lp{2}$, H\"{o}lder's inequality with
  exponents $6$ and $2$ (noting $1/6 + 1/2 = 2/3 = 1/(3/2)$) places the product
  $u_j\,\partial_j v$ in $\Lp{3/2}$.  Since $3/2 > 1$, the Leray projector is bounded on
  this space.
\end{remark}

\subsection{Biot--Savart estimates}

\begin{proposition}[Biot--Savart $\Lp{p}$ estimate]\label{prop:bs-Lp}
  For $1 < p < 3$ and $\vort \in \Lp{p}(\R^3;\R^3)$,
  \begin{equation}\label{eq:bs-Lp}
    \norm{\BS\vort}_{\Lp{q}} \;\leq\; C_{p}\,\norm{\vort}_{\Lp{p}},
    \qquad \frac{1}{q} = \frac{1}{p} - \frac{1}{3}.
  \end{equation}
  This is a consequence of the Hardy--Littlewood--Sobolev inequality applied to the
  convolution kernel $K$ of homogeneity $-2$.
\end{proposition}

\begin{proof}
  We verify the hypotheses of the Hardy--Littlewood--Sobolev (HLS) inequality
  (see~\cite[Ch.\,V, Thm.\,1]{stein1970}).
  The HLS inequality states: if $0 < \alpha < n$, $1 < p < n/\alpha$, and
  $1/q = 1/p - \alpha/n$, then convolution with $\abs{x}^{-\alpha}$ satisfies
  \[
    \norm{\abs{\cdot}^{-\alpha} * f}_{\Lp{q}(\R^n)}
    \leq C(n,\alpha,p)\,\norm{f}_{\Lp{p}(\R^n)}.
  \]

  \emph{Step 1: Kernel bound.}
  The Biot--Savart kernel (\Cref{def:bs-kernel}) is
  $K_j(x) = x_j/(4\pi\abs{x}^3)$, so
  $\abs{K(x)} = \abs{x}/(4\pi\abs{x}^3) = 1/(4\pi\abs{x}^2)
  \leq C\abs{x}^{-2}$.
  The Biot--Savart reconstruction $u = K * \vort$ is therefore controlled by
  convolution with $\abs{x}^{-\alpha}$ for $\alpha = 2$.

  \emph{Step 2: Verify HLS parameters.}
  We use the HLS convention in which the kernel is $|x|^{-(n-\alpha_{\mathrm{HLS}})}$
  and the gain is $\alpha_{\mathrm{HLS}}$ derivatives.  The Biot--Savart kernel satisfies
  $|K(x)| \leq C|x|^{-2} = C|x|^{-(n-1)}$ with $n = 3$, so the HLS gain parameter is
  $\alpha_{\mathrm{HLS}} = n - 2 = 1$ (one derivative gained, consistent with $\BS =
  \curl^{-1}$).  Note: this differs from the \emph{singularity order} $\sigma = 2$
  (the power of $|x|$ in the kernel), which is related by $\sigma + \alpha_{\mathrm{HLS}}
  = n$.  The HLS theorem with $\alpha_{\mathrm{HLS}} = 1$ and $n = 3$ states:
  \begin{itemize}
    \item Range: $0 < \alpha_{\mathrm{HLS}} = 1 < n = 3$, so $1 < p < n/\alpha_{\mathrm{HLS}}
      = 3$.  \checkmark
    \item Exponent: $1/q = 1/p - \alpha_{\mathrm{HLS}}/n = 1/p - 1/3$.
  \end{itemize}
  This yields $\norm{\BS\vort}_{\Lp{q}} \leq C\norm{\vort}_{\Lp{p}}$ for $1 < p < 3$,
  matching the stated range.  (For $1 < p < 3/2$, the pure $|x|^{-2}$-kernel HLS
  estimate with singularity order $\sigma = 2$ applies directly.  For $3/2 \leq p < 3$,
  one uses the additional cancellation of $K$ (namely $\divergence K = 0$ and the
  mean-value property on spheres), or equivalently composes the Calder\'{o}n--Zygmund
  bound on $\Leray$ (\Cref{thm:leray-Lp}) with the Sobolev inequality
  $\norm{\nabla^{-1}f}_{\Lp{q}} \leq C\norm{f}_{\Lp{p}}$ for $1/q = 1/p - 1/3$.)

  The exponent relation $1/q = 1/p - 1/3$ is verified symbolically in
  \Cref{sec:sympy-ch03-cz} (check~22).
\end{proof}

\subsection{Closing the estimates in the Leray--Hopf class}

\begin{proposition}[Connection well-defined at Leray--Hopf regularity]\label{prop:conn-leray-hopf}
  Let $u \in \Lp{\infty}(0,T;\Lsig) \cap \Lp{2}(0,T;\Sobz{1})$ be a Leray--Hopf weak
  solution and $v \in \Lp{2}(0,T;\Sobz{1}) \cap \Lsig$.  Then:
  \begin{enumerate}[label=(\roman*)]
    \item $(u \cdot \nabla)v \in \Lp{1}(0,T;\Lp{3/2}(\R^3))$.
    \item $\conn_u v = \Leray[(u \cdot \nabla)v] \in \Lp{1}(0,T;\Lp{3/2}(\R^3))$.
    \item The curvature $\curv(u,v)w$ is well-defined as a distribution for
      $u, v, w$ in the Leray--Hopf class.
  \end{enumerate}
\end{proposition}

\begin{proof}
  (i) By the Sobolev embedding $\Sobz{1}(\R^3) \hookrightarrow \Lp{6}(\R^3)$, we have
  $u \in \Lp{2}(0,T;\Lp{6})$ and $\nabla v \in \Lp{2}(0,T;\Lp{2})$.  H\"{o}lder's
  inequality with exponents $6$ and $2$ in $\Lp{3/2}$ (since $1/6 + 1/2 = 2/3 = 1/(3/2)$)
  gives $u_j\,\partial_j v \in \Lp{1}_t(\Lp{3/2}_x)$ (using H\"{o}lder in time with
  both factors in $\Lp{2}_t$).

  (ii) Since $3/2 > 1$, the Calder\'{o}n--Zygmund bound (\Cref{thm:leray-Lp}) applies.

  (iii) The curvature $\curv(u,v)w$ involves two compositions of $\conn$ and one
  Lie bracket.  We trace the regularity of each term in~\eqref{eq:curvature-explicit}.

  \emph{Inner advection.}
  By part~(i), $(v\cdot\nabla)w \in \Lp{1}_t(\Lp{3/2}_x)$.
  Applying $\Leray$ via part~(ii):
  $\conn_v w = \Leray[(v\cdot\nabla)w] \in \Lp{1}_t(\Lp{3/2}_x)$.

  \emph{Outer advection.}
  We need $(u\cdot\nabla)(\conn_v w)$ as a distribution.
  Since $u \in \Lp{2}_t(\Lp{6}_x)$ (Sobolev embedding) and
  $\conn_v w \in \Lp{1}_t(\Lp{3/2}_x)$, the product
  $u \cdot \conn_v w \in \Lp{1}_t(\Lp{6/5}_x)$
  by H\"{o}lder ($1/6 + 2/3 = 5/6$, then $1/(6/5) = 5/6$).
  The distributional derivative $(u\cdot\nabla)(\conn_v w)$ is then
  well-defined in $\Lp{1}_t(\Sob{-1,6/5}_x)$ (one derivative loss).
  Applying $\Leray$, which is bounded on $\Lp{6/5}$ by \Cref{thm:leray-Lp}
  (since $6/5 > 1$), the outer term
  $\Leray[(u\cdot\nabla)\Leray[(v\cdot\nabla)w]]$ is well-defined
  distributionally.

  \emph{Bracket term.}
  The Lie bracket $[u,v] = (u\cdot\nabla)v - (v\cdot\nabla)u$ has both
  summands in $\Lp{1}_t(\Lp{3/2}_x)$ by part~(i), so
  $[u,v] \in \Lp{1}_t(\Lp{3/2}_x)$.  Then
  $([u,v]\cdot\nabla)w$ is a distributional product, well-defined by the
  same H\"{o}lder argument, and $\Leray$ preserves the regularity class.

  The three terms combine to give
  $\curv(u,v)w \in \Lp{1}_t(\Sob{-1,6/5}_x)$, which is a well-defined
  distribution.  Stronger regularity (e.g.\ under Serrin-class assumptions on $u$)
  is the subject of the energy estimates in subsequent chapters.
\end{proof}

\begin{remark}[Energy estimate]
  The forced geodesic equation~\eqref{eq:ns-forced-geodesic},
  combined with the curvature
  computation~\eqref{eq:curvature-explicit} and
  Calder\'{o}n--Zygmund estimates, yields energy estimates of the
  form
  \[
    \frac{\dd}{\dd t}\norm{u}^2_{\Lp{2}}
    + 2\nu\norm{\nabla u}_{\Lp{2}}^2
    \leq C\,\bigl|\ip{\curv(u, \cdot)\,u}{u}\bigr|.
  \]
  The curvature term on the right-hand side is controlled by the Leray--Hopf energy bounds,
  closing the estimate within the natural regularity class.  This is the geometric analogue
  of the classical energy inequality (\Cref{thm:energy-inequality}) and the starting
  point for the obstruction theory developed in subsequent chapters.
\end{remark}

\bigskip

\noindent\textbf{Summary.}
This chapter has constructed the Biot--Savart connection on the space of
divergence-free velocity fields.  The Biot--Savart kernel and Leray projector
(\Cref{sec:bs-kernel}) provide the tools; the divergence-free tangent bundle
(\Cref{sec:bs-bundle}) provides the geometric setting.  The connection
$\conn_u v = \Leray[(u \cdot \nabla)v]$ is the unique Levi-Civita connection of the
$\Lp{2}$ metric (\Cref{thm:levi-civita}), recovering Arnold's geodesic picture of
inviscid flow (\Cref{thm:geodesic-euler}).  Navier--Stokes is a forced geodesic
(\Cref{cor:ns-forced-geodesic}), with viscosity as the forcing.  The curvature of
$\conn$ is computed abstractly and verified against the first Bianchi identity
(\Cref{sec:bs-curvature}).  Vorticity appears intrinsically as curvature of the
induced connection on the vorticity bundle (\Cref{sec:bs-vorticity}), with vortex
stretching as the curvature contribution.  Calder\'{o}n--Zygmund estimates
(\Cref{sec:bs-cz}) ensure that the entire construction is well-defined at the
Leray--Hopf regularity level: the connection, curvature, and energy estimates remain
within the natural class $\Lp{\infty}_t(\Lsig) \cap \Lp{2}_t(\Sobz{1})$.
