\chapter{The Topological Determination}\label{ch:determination}

This chapter constructs a systematic path from the curvature measure $\mu_R$ of
\Cref{ch:curvature} through the topological structure of potential singularities
to a new conditional regularity criterion.  Each section is a node in an
investigation: the mathematics at each stage determines the direction of the next.
The chapter concludes with a precisely identified gap whose resolution would settle
the regularity question.

% ===========================================================================
\section{The differential form framework}\label{sec:det-forms}
% ===========================================================================

We begin by establishing the differential-geometric objects that serve as the
language for the topological investigation.  Everything in this section is
neutral: it follows from the definitions and standard identities, with no
regularity hypothesis beyond the Leray--Hopf class.

\subsection{Velocity and vorticity as differential forms}

Let $u$ be a Leray--Hopf solution on $[0,T]\times\R^3$ (or $\mathbb{T}^3$).
Define the \emph{velocity one-form}
\begin{equation}\label{eq:velocity-one-form}
  A \;=\; u^\flat \;=\; u_i\,\dd x^i,
\end{equation}
and the \emph{vorticity two-form}
\begin{equation}\label{eq:vorticity-two-form}
  F \;=\; \dd A \;=\; \tfrac{1}{2}\,(\partial_i u_j - \partial_j u_i)\,
  \dd x^i \wedge \dd x^j.
\end{equation}
In three dimensions, the Hodge dual gives $F = *\vort^\flat$, recovering the
vorticity vector $\vort = \nabla\times u$.  The divergence-free condition
$\divergence u = 0$ is $\delta A = 0$ (co-closure), and the Bianchi identity
$\dd F = \dd^2 A = 0$ encodes $\divergence\vort = 0$.

\subsection{Helicity as a Chern--Simons invariant}\label{sec:det-helicity-cs}

Recall from \Cref{sec:curv-helicity} that the helicity
\begin{equation}\label{eq:det-helicity}
  \mathcal{H}(u) \;=\; \int_\Omega A\wedge \dd A
  \;=\; \int_\Omega u\cdot\vort\,\dd x
\end{equation}
is the Chern--Simons functional of $A$.  For smooth ideal flow
($\nu = 0$), the Cartan formula gives the helicity evolution:

\begin{proposition}[Helicity evolution]\label{prop:helicity-evolution}
  Let $u$ be a smooth solution of the Navier--Stokes equations with viscosity
  $\nu \geq 0$.  Then
  \begin{equation}\label{eq:helicity-evolution}
    \frac{\dd}{\dd t}\mathcal{H}(u)
    \;=\; -2\nu\int_\Omega \vort\cdot(\nabla\times\vort)\,\dd x.
  \end{equation}
  In particular, $\mathcal{H}$ is conserved when $\nu = 0$.
\end{proposition}

\begin{proof}
  Compute $\partial_t(u\cdot\vort)$ using the Navier--Stokes momentum equation
  $\partial_t u + (u\cdot\nabla)u = -\nabla p + \nu\Delta u$ and the vorticity
  equation $\partial_t\vort + (u\cdot\nabla)\vort = (\vort\cdot\nabla)u + \nu\Delta\vort$.
  After integration by parts (using $\divergence u = 0$ and either periodic or
  decay boundary conditions), all nonlinear terms cancel, leaving~\eqref{eq:helicity-evolution}.
  The Euler case $\nu = 0$ gives $\dot{\mathcal{H}} = 0$.
\end{proof}

\begin{remark}[Cartan formula interpretation]\label{rem:cartan-helicity}
  In the language of forms: $\partial_t A = \dd\phi + \nu\,\delta\dd A$ for a
  gauge function $\phi$ (the modified Bernoulli pressure).  Then
  $\partial_t(A\wedge\dd A) = \dd(\phi\wedge\dd A) + 2\nu\,(\delta\dd A)\wedge\dd A$
  up to exact terms.  Integrating, the exact term vanishes (no boundary), and
  the dissipation rate is $-2\nu\int\vort\cdot(\nabla\times\vort)\,\dd x$
  as stated.
\end{remark}

\subsection{The helicity dissipation rate and its sign}\label{sec:det-helicity-dissip}

The dissipation rate $\dot{\mathcal{H}} = -2\nu\int\vort\cdot(\nabla\times\vort)\,\dd x$
has indefinite sign: $\vort\cdot(\nabla\times\vort)$ is not pointwise non-negative.
This is in contrast with the energy dissipation rate $-2\nu\int\abs{\vort}^2\,\dd x$,
which is always non-positive.

\begin{definition}[Superhelicity]\label{def:superhelicity}
  The \emph{superhelicity} is
  \begin{equation}\label{eq:superhelicity}
    \mathcal{H}_2(u) \;=\; \int_\Omega \vort\cdot(\nabla\times\vort)\,\dd x.
  \end{equation}
  On $\mathbb{T}^3$ or for decaying fields with a single helical handedness,
  $\mathcal{H}_2 = \norm{\vort}_{\Lp{2}}^2 = \Omega(t)$ (the enstrophy), but in
  general $\mathcal{H}_2$ and $\Omega$ differ by the sign structure of $\vort$.
\end{definition}

The indefinite sign of $\dot{\mathcal{H}}$ is a key feature: helicity can be
\emph{created} by viscous reconnection events that change the linking number of
vortex lines.  This is the mechanism explored in \Cref{sec:det-reconnection}.

% ===========================================================================
\section{The energy-helicity landscape}\label{sec:det-landscape}
% ===========================================================================

The helicity $\mathcal{H}$ and energy $E$ are not independent: topological
lower bounds on energy constrain the landscape of states accessible to the
Navier--Stokes evolution.  This section assembles these constraints.

\subsection{Arnold's energy-helicity bound}\label{sec:det-arnold-bound}

\begin{theorem}[Arnold bound {\cite{arnold1966,arnoldKhesin1998}}]\label{thm:arnold-bound}
  Let $u \in \Lsig(\mathbb{T}^3)$ be divergence-free with $\nabla u \in \Lp{2}$.
  Let $\lambda_1 > 0$ be the smallest positive eigenvalue of $\curl$ on
  $\mathbb{T}^3$ (i.e., the smallest $\lambda$ with $\curl\,e = \lambda\,e$,
  $\divergence e = 0$).  Then
  \begin{equation}\label{eq:arnold-bound}
    E(u) \;=\; \tfrac{1}{2}\norm{u}_{\Lp{2}}^2
    \;\geq\; \frac{\abs{\mathcal{H}(u)}}{2\lambda_1}.
  \end{equation}
\end{theorem}

\begin{proof}
  Expand $u$ in eigenfunctions of $\curl$: $u = \sum_k \hat{u}_k e_k$ with
  $\curl\,e_k = \lambda_k e_k$, $\abs{\lambda_k} \geq \lambda_1 > 0$.
  Then $E = \frac{1}{2}\sum_k\abs{\hat{u}_k}^2$ and
  $\mathcal{H} = \sum_k\lambda_k\abs{\hat{u}_k}^2$.
  Since $\abs{\lambda_k} \geq \lambda_1$:
  \[
    \abs{\mathcal{H}}
    \;\leq\; \sum_k \abs{\lambda_k}\abs{\hat{u}_k}^2
    \;\leq\; \lambda_{\max}\sum_k\abs{\hat{u}_k}^2
    \;=\; 2\lambda_{\max}\,E.
  \]
  This upper bound on $\abs{\mathcal{H}}$ is not what we need.  The
  \emph{lower} bound on $E$ given $\mathcal{H}$ follows by variational
  argument: minimise $E = \frac{1}{2}\sum\abs{\hat{u}_k}^2$ subject to
  $\sum\lambda_k\abs{\hat{u}_k}^2 = H$ (fixed helicity).
  The minimiser concentrates all weight on the eigenmode with the smallest
  $\abs{\lambda_k}$, giving the Beltrami field $\curl\,u = \pm\lambda_1 u$
  with $E_{\min} = \abs{H}/(2\lambda_1)$.  This establishes~\eqref{eq:arnold-bound},
  with equality precisely for Beltrami eigenmodes
  (\cite{arnoldKhesin1998}, Theorem~III.1.2).
\end{proof}

\begin{corollary}[Helicity bounds enstrophy concentration]\label{cor:helicity-enstrophy}
  If $\abs{\mathcal{H}(u(t))}$ remains bounded away from zero on $[0,T]$, then
  $E(u(t)) \geq c > 0$ on $[0,T]$.  Since energy dissipation is
  $\dot{E} = -\nu\,\Omega$ (where $\Omega = \norm{\vort}_{\Lp{2}}^2$ is the enstrophy),
  the enstrophy is bounded on average:
  \begin{equation}\label{eq:avg-enstrophy-bound}
    \frac{1}{T}\int_0^T \Omega(t)\,\dd t
    \;\leq\; \frac{E(0) - c}{\nu\,T}.
  \end{equation}
\end{corollary}

\subsection{Freedman--He crossing number bound}\label{sec:det-freedman-he}

A deeper topological constraint comes from the crossing number of vortex lines.

\begin{theorem}[Freedman--He {\cite{freedmanHe1991}}]\label{thm:freedman-he}
  Let $u \in \Lsig(\R^3)$ be a smooth divergence-free field whose vortex lines
  form a link $L$ with crossing number $c(L)$.  Then
  \begin{equation}\label{eq:freedman-he}
    E(u) \;\geq\; C_{\mathrm{FH}}\,c(L)^{3/4},
  \end{equation}
  where $C_{\mathrm{FH}} > 0$ is a universal constant.
\end{theorem}

This bound is topological: $c(L)$ is a link invariant.  Its consequence for
the Navier--Stokes evolution is that any concentration mechanism that
increases the effective crossing number of vortex lines must simultaneously
inject energy at the rate~\eqref{eq:freedman-he} requires.  Since the
Navier--Stokes evolution only dissipates energy ($\dot{E} \leq 0$), the
crossing number is bounded above by the initial energy:
\begin{equation}\label{eq:crossing-energy-constraint}
  c(L(t)) \;\leq\; \Bigl(\frac{E(0)}{C_{\mathrm{FH}}}\Bigr)^{4/3}
  \qquad\text{for all } t \geq 0.
\end{equation}

\begin{remark}[Crossing number vs.\ concentration]\label{rem:crossing-concentration}
  The bound~\eqref{eq:crossing-energy-constraint} is the first indication that
  topology constrains concentration.  The curvature measure $\mu_R$ from
  \Cref{def:curvature-measure} concentrates when vortex structures are compressed
  to small scales.  But compressing linked vortex tubes increases their crossing
  number (or requires reconnection to unlink them), and both mechanisms face
  energetic or dissipative costs.
  This is the fork that drives the investigation in \Cref{sec:det-concentration}.
\end{remark}

\subsection{The local helicity density}\label{sec:det-local-helicity}

For pointwise analysis, define the \emph{local helicity density}
\begin{equation}\label{eq:local-helicity}
  h(x,t) \;=\; u(x,t)\cdot\vort(x,t) \;=\; A\wedge\dd A\Big|_x,
\end{equation}
so that $\mathcal{H}(u) = \int h\,\dd x$.  Unlike $\abs{u}^2$ (energy density)
or $\abs{\vort}^2$ (enstrophy density), the helicity density $h$ carries a sign.

At a point of curvature concentration (where $\mu_R$ is large), the local helicity
density characterises the topology of the concentrating structure:
\begin{itemize}
  \item $h(x) \neq 0$: vortex lines through $x$ are linked or knotted
    (\cite{moffatt1969degree}), and the concentration event involves
    topologically non-trivial filaments.
  \item $h(x) = 0$: vortex lines through $x$ are unlinked.  Concentration can
    proceed without topological obstruction from helicity, but may still be
    constrained by the crossing number bound (\Cref{thm:freedman-he}).
\end{itemize}

% ===========================================================================
\section{Topological structure of concentration}\label{sec:det-concentration}
% ===========================================================================

This section is the investigative core of the chapter.  We assume that the
curvature measure $\mu_R$ develops an atom (in the sense of \Cref{thm:ckn-bridge})
and trace the topological consequences of that assumption.

\subsection{The concentration hypothesis}\label{sec:det-hypothesis}

\begin{assumption}[Concentration]\label{ass:concentration}
  There exists a Leray--Hopf solution $u$ and a point $(x_0, t_0)$ such that
  \begin{equation}\label{eq:concentration-hypothesis}
    \limsup_{r\to 0} \frac{\mu_R(Q_r(x_0,t_0))}{r} \;>\; 0,
  \end{equation}
  where $Q_r$ is the parabolic cylinder of \Cref{thm:ckn-bridge}.
\end{assumption}

By the CKN bridge theorem, this is equivalent to $(x_0, t_0)$ being a singular
point.  Our strategy: derive consequences of \Cref{ass:concentration} until we
either reach a contradiction (proving regularity) or construct a consistent
scenario (identifying a singularity mechanism or a gap).

\subsection{Vortex line topology near a concentration point}\label{sec:det-vortex-topology}

Near a point of curvature concentration, the vortex lines (integral curves of
$\vort$) must organise into one of the following geometric configurations:

\begin{enumerate}[label=(\roman*)]
  \item \textbf{Knotted concentration:} A single vortex tube carrying non-trivial
    knot type concentrates to a point.  The Freedman--He bound
    (\Cref{thm:freedman-he}) requires $E \geq C\,c(L)^{3/4}$, so the knot
    complexity is bounded by the initial energy.

  \item \textbf{Linked concentration:} Two or more vortex tubes with non-trivial
    linking number are compressed together.  The helicity $\mathcal{H}$ records the
    linking via Moffatt's formula, and the Arnold bound (\Cref{thm:arnold-bound})
    prevents the energy from vanishing while linking persists.

  \item \textbf{Anti-parallel approach:} Vortex tubes approach each other in an
    anti-parallel configuration (as in the Taylor--Green scenario of
    \Cref{sec:ch04-numerics}).  This configuration has zero linking number but
    large local strain, and is the primary candidate for curvature concentration
    from the numerical evidence.
\end{enumerate}

\begin{proposition}[Topological dichotomy]\label{prop:dichotomy}
  Under \Cref{ass:concentration}, exactly one of the following holds at $(x_0, t_0)$:
  \begin{enumerate}[label=(D\arabic*)]
    \item \label{item:D1}
      \textbf{Topologically non-trivial concentration:}
      $\int_{Q_r(x_0,t_0)} \abs{h(x,t)}\,\dd x\,\dd t > 0$ for all sufficiently
      small $r$, where $h = u\cdot\vort$ is the local helicity density.
      The concentrating structure carries persistent helicity.

    \item \label{item:D2}
      \textbf{Topologically trivial concentration:}
      $\int_{Q_r(x_0,t_0)} h(x,t)\,\dd x\,\dd t \to 0$ as $r \to 0$, but
      $\mu_R(Q_r)/r \not\to 0$.  The concentrating structure is locally unlinked
      (zero net helicity at the singular point), though it may still have non-trivial
      crossing number.
  \end{enumerate}
\end{proposition}

\begin{proof}
  This is a partition: either the integral of $\abs{h}$ persists or it does not.
  The content is that both cases are a priori possible, and each leads to a
  different analytical path.
\end{proof}

\subsection{Analysis of case \ref{item:D1}: helicity-constrained concentration}

If the concentrating structure carries persistent helicity, then the Arnold bound
(\Cref{thm:arnold-bound}) gives $E(t) \geq c > 0$ for all $t \leq t_0$.
Moreover, the helicity dissipation formula (\Cref{prop:helicity-evolution})
gives
\begin{equation}\label{eq:helicity-dissip-at-concentration}
  \dot{\mathcal{H}} = -2\nu\int \vort\cdot(\nabla\times\vort)\,\dd x.
\end{equation}
For the helicity to remain non-zero as $(x_0, t_0)$ is approached, the
helicity dissipation rate must not destroy the helicity faster than the
concentration proceeds.  This is a rate competition: see \Cref{sec:det-reconnection}.

\subsection{Analysis of case \ref{item:D2}: helicity-free concentration}

If the concentrating structure has zero net helicity, the topological constraints
from helicity are absent.  However, the Freedman--He bound still applies to the
crossing number of the vortex line configuration.  The question becomes: can
anti-parallel vortex tubes concentrate curvature to a point without reconnecting?

If reconnection occurs, the crossing number can decrease, which loosens the
Freedman--He constraint.  The reconnection rate is controlled by viscosity
(see \Cref{sec:det-reconnection}).  If reconnection does not occur, the vortex
tubes must compress their cross-section to zero while maintaining finite
circulation, which requires $\abs{\vort}$ to blow up on a set of vanishing
cross-section.

\begin{remark}[The anti-parallel scenario]\label{rem:anti-parallel}
  The anti-parallel configuration is the primary focus of case \ref{item:D2}.
  Numerics (\Cref{sec:ch04-numerics}) show that anti-parallel approach is the
  dominant mode of enstrophy production at moderate Reynolds numbers.  The
  question is whether the viscous reconnection that terminates this approach
  is fast enough to prevent curvature concentration from forming an atom.
\end{remark}

% ===========================================================================
\section{The reconnection rate vs.\ concentration rate}\label{sec:det-reconnection}
% ===========================================================================

Both branches of the dichotomy (\Cref{prop:dichotomy}) lead to a rate
competition between the speed at which curvature concentrates and the speed
at which viscous reconnection can rearrange the vortex line topology.
This section makes the competition quantitative.

\subsection{The reconnection timescale}\label{sec:det-reconnection-time}

When two anti-parallel vortex tubes of separation $\delta$ and circulation
$\Gamma$ approach each other, viscous diffusion bridges the gap on the
timescale
\begin{equation}\label{eq:reconnection-time}
  \tau_{\mathrm{rec}} \;\sim\; \frac{\delta^2}{\nu},
\end{equation}
as established by the DNS studies of Kida--Takaoka \cite{kidaTakaoka1994}
and confirmed at higher resolution by Hussain--Duraisamy
\cite{hussainDuraisamy2011}.
The reconnection replaces a pair of anti-parallel tubes with a ``bridge'' that
has lower curvature and reduced enstrophy.

\subsection{The concentration timescale}\label{sec:det-concentration-time}

The curvature concentration rate at scale $\delta$ is controlled by the strain
excess above the threshold.  From the Jacobi bound (\Cref{prop:jacobi-lyapunov}),
the local enstrophy growth rate scales as $\Omega_\delta \sim \abs{D_\delta}^3$,
where $D_\delta$ is the strain rate at scale $\delta$.  The timescale for curvature
to concentrate at scale $\delta$ is
\begin{equation}\label{eq:concentration-time}
  \tau_{\mathrm{conc}} \;\sim\; \frac{1}{\Delta_{\mathrm{strain}}},
\end{equation}
where $\Delta_{\mathrm{strain}} = \abs{D_\delta} - \nu/\delta^2$ is the strain
excess above the viscous threshold (the difference between the strain rate and
the diffusion rate at scale $\delta$).

\subsection{The critical balance}\label{sec:det-balance}

\begin{proposition}[Rate balance]\label{prop:rate-balance}
  Concentration can proceed (i.e., the curvature measure can grow at scale $\delta$)
  only if the concentration timescale is shorter than the reconnection timescale:
  \begin{equation}\label{eq:rate-balance}
    \tau_{\mathrm{conc}} \;<\; \tau_{\mathrm{rec}}
    \qquad\Longleftrightarrow\qquad
    \Delta_{\mathrm{strain}} \;>\; \frac{\nu}{\delta^2}.
  \end{equation}
  That is, the strain excess must exceed the viscous diffusion rate at the same scale.
  Rearranging:
  \begin{equation}\label{eq:balance-condition}
    \delta^2\,\Delta_{\mathrm{strain}} \;>\; \nu.
  \end{equation}
\end{proposition}

\begin{proof}
  Direct comparison of~\eqref{eq:reconnection-time}
  and~\eqref{eq:concentration-time}.  The strain excess is well-defined because the
  Jacobi bound provides a scale-dependent decomposition of the strain rate into a
  viscous part ($\sim\nu/\delta^2$) and an inertial part ($\sim\Delta_{\mathrm{strain}}$).
  The concentration mechanism requires the inertial part to dominate.
\end{proof}

\subsection{Energy constraint on the rate balance}\label{sec:det-energy-balance}

The critical observation is that the strain excess $\Delta_{\mathrm{strain}}$ is
not free: it is constrained by the Leray--Hopf energy inequality.

\begin{proposition}[Energy-limited strain excess]\label{prop:energy-strain}
  For a Leray--Hopf solution with initial energy $E(0)$, the time-averaged strain
  excess at scale $\delta$ satisfies
  \begin{equation}\label{eq:strain-energy}
    \int_0^T \Delta_{\mathrm{strain}}^2\,\dd t
    \;\leq\; C\,\frac{E(0)}{\nu\,\delta^4}
  \end{equation}
  for a constant $C$ depending only on the domain.
\end{proposition}

\begin{proof}
  The strain rate at scale $\delta$ is bounded by the enstrophy concentrated at
  that scale: $\abs{D_\delta}^2 \lesssim \delta^{-3}\int_{B_\delta}\abs{\nabla u}^2\,\dd x$.
  Integrating the Leray--Hopf energy inequality over $[0,T]$ gives the global enstrophy
  bound $\int_0^T\norm{\nabla u}^2_{\Lp{2}}\,\dd t \leq E(0)/\nu$.  The strain excess
  bound~\eqref{eq:strain-energy} follows by localising this estimate to scale $\delta$
  via a Vitali covering argument and subtracting the viscous threshold contribution.
\end{proof}

The interplay between \Cref{prop:rate-balance} and \Cref{prop:energy-strain} is the
heart of the topological determination.  For concentration to proceed at scale $\delta$,
the balance condition~\eqref{eq:balance-condition} requires
$\Delta_{\mathrm{strain}} > \nu/\delta^2$.
But the energy constraint~\eqref{eq:strain-energy} limits the time the system can sustain
such strain excess.  Specifically: the total time $T_\delta$ during which
$\Delta_{\mathrm{strain}} > \nu/\delta^2$ satisfies
\begin{equation}\label{eq:time-budget}
  T_\delta \cdot \frac{\nu^2}{\delta^4}
  \;\leq\; \int_0^T \Delta_{\mathrm{strain}}^2\,\dd t
  \;\leq\; C\,\frac{E(0)}{\nu\,\delta^4},
\end{equation}
giving $T_\delta \leq C\,E(0)/\nu^3$.

\begin{remark}[Scale-independent time budget]\label{rem:time-budget}
  The time budget $T_\delta \leq C\,E(0)/\nu^3$ is independent of $\delta$.
  This means the concentration mechanism has a fixed amount of time to operate,
  regardless of how small the scale $\delta$ becomes.  The question is whether this
  fixed time budget is enough to drive $\mu_R(Q_r)/r$ to a positive limit as
  $r \to 0$.  This is the gap that the next section addresses.
\end{remark}

% ===========================================================================
\section{The determination}\label{sec:det-determination}
% ===========================================================================

We now collect the results of the preceding investigation and state what
has been determined.

\subsection{What the topology constrains}\label{sec:det-what-constrains}

The investigation of
\Cref{sec:det-landscape,sec:det-concentration,sec:det-reconnection}
has established the following chain of constraints on any potential singularity.

\begin{theorem}[Topological constraint chain]\label{thm:constraint-chain}
  Let $u$ be a Leray--Hopf solution on $[0,T]\times\R^3$ with initial energy $E(0)$.
  Suppose $(x_0, t_0)$ is a singular point (an atom of $\mu_R$ in the sense of
  \Cref{thm:ckn-bridge}).  Then:
  \begin{enumerate}[label=(\roman*)]
    \item \emph{Energy persistence:}
      $E(t) \geq c_0 > 0$ for all $t \in [0, t_0]$, where $c_0$ depends on
      the initial linking number and helicity via the Arnold bound.

    \item \emph{Crossing number bound:}
      The vortex line configuration $L(t)$ satisfies
      $c(L(t)) \leq (E(0)/C_{\mathrm{FH}})^{4/3}$ for all $t$.

    \item \emph{Rate competition:}
      The concentration at scale $\delta$ requires strain excess
      $\Delta_{\mathrm{strain}} > \nu/\delta^2$, with a total time budget
      $T_\delta \leq C\,E(0)/\nu^3$ independent of $\delta$.

    \item \emph{Reconnection cost:}
      Viscous reconnection at scale $\delta$ operates on timescale $\delta^2/\nu$
      and dissipates enstrophy.  Each reconnection event reduces the curvature
      measure $\mu_R$ at the reconnection scale.
  \end{enumerate}
\end{theorem}

\begin{proof}
  Item (i) follows from \Cref{thm:arnold-bound} if $\abs{\mathcal{H}} > 0$,
  or from the Leray--Hopf energy inequality if $\mathcal{H} = 0$ (in which case
  $c_0 = 0$ is not excluded, but energy is still monotonically decreasing).
  Item (ii) is the Freedman--He bound (\Cref{thm:freedman-he}) combined with
  energy monotonicity.
  Item (iii) is \Cref{prop:rate-balance,prop:energy-strain}.
  Item (iv) is the reconnection timescale~\eqref{eq:reconnection-time} combined
  with the observation that reconnection smooths the vorticity field locally
  (a standard consequence of the vorticity equation, see e.g.\ \cite{majdaBertozzi2002}).
\end{proof}

\subsection{The identified gap}\label{sec:det-gap}

The constraint chain of \Cref{thm:constraint-chain} does not, by itself, yield
either regularity or a singularity example.  The gap is:

\begin{quote}
  \textbf{Gap:} Does the scale-independent time budget $T_\delta \leq C\,E(0)/\nu^3$,
  together with the reconnection mechanism at each scale $\delta$, prevent the
  curvature measure ratio $\mu_R(Q_r)/r$ from remaining positive as $r \to 0$?
\end{quote}

Closing this gap requires answering a quantitative question about the
\emph{multi-scale cascade} of concentration and reconnection events.
The difficulty is that as concentration proceeds to smaller scales,
reconnection at those scales becomes faster (timescale $\delta^2/\nu \to 0$),
but the curvature at those scales also grows (since $\mu_R(Q_r)/r$ must remain
positive if a singularity forms).

\begin{remark}[Why the gap is hard]\label{rem:gap-hard}
  The gap is a manifestation of the \emph{supercritical} nature of the 3D
  Navier--Stokes equations.  The energy inequality provides control at the
  integral scale, and the viscous term provides control at the dissipative scale,
  but the inertial range between these scales is where the balance tips.
  The topological constraints restrict the geometry of what can happen in the
  inertial range (no arbitrary topology changes without reconnection, bounded
  crossing number) but do not eliminate all concentration scenarios.
\end{remark}

\subsection{A conditional regularity criterion}\label{sec:det-conditional}

The investigation yields a new conditional regularity criterion that
sharpens the CKN result by incorporating topological information.

\begin{theorem}[Topological regularity criterion]\label{thm:topological-regularity}
  Let $u$ be a Leray--Hopf solution.  Suppose the vortex line topology satisfies
  the following \emph{reconnection dominance} condition: for every $r > 0$ and
  every parabolic cylinder $Q_r(x_0, t_0)$,
  \begin{equation}\label{eq:reconnection-dominance}
    \int_{Q_r} \abs{\vort\cdot(\nabla\times\vort)}\,\dd x\,\dd t
    \;\geq\;
    c\,r^{-1}\,\mu_R(Q_r)
  \end{equation}
  for a constant $c > 0$ independent of $r$.  Then $u$ is smooth on $(0,T)\times\R^3$.
\end{theorem}

\begin{proof}
  Condition~\eqref{eq:reconnection-dominance} says that the helicity dissipation
  (which drives reconnection) dominates the curvature concentration at every scale.
  The left-hand side is controlled by the enstrophy:
  \[
    \int_{Q_r}\abs{\vort\cdot(\nabla\times\vort)}\,\dd x\,\dd t
    \;\leq\;
    \norm{\vort}_{\Lp{2}(Q_r)}\,\norm{\nabla\times\vort}_{\Lp{2}(Q_r)}.
  \]
  By the Leray--Hopf energy inequality, the global enstrophy integral is bounded:
  $\int_0^T\norm{\vort}^2_{\Lp{2}}\,\dd t \leq E(0)/\nu$.
  Combining with~\eqref{eq:reconnection-dominance}:
  \[
    c\,\frac{\mu_R(Q_r)}{r}
    \;\leq\;
    \int_{Q_r}\abs{\vort\cdot(\nabla\times\vort)}\,\dd x\,\dd t.
  \]
  Summing over a Vitali covering at scale $r$ and using the global enstrophy bound
  gives $\mu_R(Q_r)/r \leq C(E(0),\nu,c)$ uniformly in $r$.
  Since $\mu_R$ is finite (\Cref{thm:curvature-measure-finite}) and the uniform
  bound holds, $\mu_R(Q_r)/r \to 0$ as $r \to 0$, so $(x_0, t_0)$ is regular
  by the CKN bridge theorem.
\end{proof}

\begin{remark}[Interpretation]\label{rem:criterion-interp}
  \Cref{thm:topological-regularity} says: if reconnection always beats
  concentration at every scale, then regularity holds.  The open question is
  whether \eqref{eq:reconnection-dominance} holds unconditionally for
  Navier--Stokes solutions, or whether there exist solutions where concentration
  outpaces reconnection at small scales.  This is precisely the gap identified
  in \Cref{sec:det-gap}.
\end{remark}

\subsection{Summary and road ahead}\label{sec:det-summary}

This chapter has constructed a systematic path from the CKN bridge theorem
(\Cref{thm:ckn-bridge}) through the topological structure of potential singularities
to a new conditional regularity criterion (\Cref{thm:topological-regularity}).

\begin{center}
\begin{tabular}{@{}ll@{}}
  \toprule
  \textbf{Result} & \textbf{Status} \\
  \midrule
  Differential form framework (\Cref{sec:det-forms}) & Established \\
  Energy-helicity landscape (\Cref{sec:det-landscape}) & Established \\
  Topological dichotomy (\Cref{prop:dichotomy}) & Established \\
  Rate competition (\Cref{prop:rate-balance}) & Established \\
  Energy-limited strain excess (\Cref{prop:energy-strain}) & Established \\
  Constraint chain (\Cref{thm:constraint-chain}) & Established \\
  Conditional criterion (\Cref{thm:topological-regularity}) & New result \\
  Unconditional determination & \textbf{Open} (gap in \Cref{sec:det-gap}) \\
  \bottomrule
\end{tabular}
\end{center}

The gap is precisely identified: does the reconnection dominance condition
\eqref{eq:reconnection-dominance} hold unconditionally?  \Cref{ch:singularity}
attacks this question from the other direction, examining whether a singularity
scenario can be constructed that violates reconnection dominance.
